---
title: "00_mixing_experiment_c-figures"
author: "Mark Robinson, Izaskun Mallona & Giulia Moro"
format: 
  html:
    toc: true
    embed-resources: true
    keep_md: true
editor_options: 
  chunk_output_type: console
---

### TO DO 

```{r}

# Figure out Fuzznuc ROI error
# Remove regulatory_region_egfp

```

### setting up knitr 

```{r}

knitr::opts_chunk$set(fig.width = 5,
               fig.height = 5,
               cache = TRUE,
               include = TRUE,
               dev = c("png","svg"),
               cache.lazy = FALSE,
               warning = TRUE,
               message = TRUE)

```


### Load packages, load data 

```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(SingleCellExperiment)
library(scran)
library(ggplot2)
library(dplyr)
library(scater)
library(reshape2)
library(UpSetR)
library(limma)
library(cowplot)
library(pheatmap)
library(readr)
library(plotROC)
library(here)
library(RColorBrewer)
library(ggnewscale)
library(scales)
library(rmarkdown)
library(irlba)
library(viridis)
library(gridExtra)
library(googlesheets4)
library(tools)

#packageVersion("Matrix")

getwd() # after opening the .Rproj file, getwd() should be "/Users/gimoro/figure_making_paper/rock_roi_paper"

rdss <- dir(here("00_mixing_experiment"), "^mixing.*rds$", 
            recursive = TRUE, full.names = TRUE)

names(rdss) <- gsub("_sce.rds", "", basename(rdss))
rdss # prints the .rds files in the 00_mixing_experiment folder

md5sum(rdss)

datadir <- dirname(rdss)[1] # folder where the .rds files are 

sces <- mapply(function(u,v)  {
  rds <- readRDS(u)
  rds$sample_id <- v
  rds$sample_id.barcode <- paste0(v,".",colnames(rds))
  # colnames(rowData(rds)) <- c("name", "type", "value","captured")
  g <- grepl("^ENS", rownames(rds))
  rownames(rds)[g] <- paste0(rownames(rds)[g], 
                             "__", rowData(rds)$name[g])
  rowData(rds)$type <- "capture"
  k <- grepl("^ENSG", rownames(rds))
  rowData(rds)$type[k] <- "human_gene"
  k <- grepl("^ENSMUS", rownames(rds))
  rowData(rds)$type[k] <- "mouse_gene"
  k <- grepl("^mt", rowData(rds)$name)
  rowData(rds)$type[k] <- "mouse_mito"
  k <- grepl("^MT", rowData(rds)$name)
  rowData(rds)$type[k] <- "human_mito"
  rds
}, rdss, names(rdss))

sce <- Reduce(cbind, sces)

```

### First figure: mito vs total UMI 

```{r}

rowData(sce)$type %>% table

rd <- rowData(sce)
sce <- addPerCellQCMetrics(sce, assay.type = "wta",
                           subsets=list(capture=rd$type=="capture",
                                        human_gene=rd$type=="human_gene",
                                        human_mito=rd$type=="human_mito",
                                        mouse_gene=rd$type=="mouse_gene",
                                        mouse_mito=rd$type=="mouse_mito"))

cd <- colData(sce) %>% as.data.frame %>% 
  mutate(total_mito_percent = subsets_mouse_mito_percent+subsets_human_mito_percent)

ggplot(cd, aes(x = total, y = total_mito_percent)) +
  geom_point() + scale_x_log10() + scale_y_sqrt() +
  facet_wrap(~factor(sample_id,levels=c("mixing_unmod","mixing_unmod_roi","mixing_rock","mixing_rockroi"))) + 
  geom_hline(yintercept=c(3,25), colour="orange") +
  geom_vline(xintercept=c(2750), colour="orange") +
  geom_density2d()+
  theme_bw()+
  xlab("total number of transcripts")+
  ylab("percent mitochondrial reads")+
  theme(axis.text = element_text(size = 15), axis.title=element_text(size=20,face="bold"),strip.text=element_text(size=15,face="bold"))

```

### Generating filtered dataset

```{r}

mito <- sce$subsets_mouse_mito_percent+sce$subsets_human_mito_percent
sce_filt <- sce[,sce$total>2750 & mito>3 & mito<25]
sce_filt <- sce_filt[rowSums(assay(sce_filt,1))>2 | rowData(sce_filt)$type=="capture" ,] 

rm(sce,sces)

```

### Defining regions in filt dataset for tso_ontarget_multi and wta_ontarget_multi

```{r}

# read annotation info from GS
anno_url <- "https://docs.google.com/spreadsheets/d/1cvldF_VFA7FYuChoR9a4SIpYcZ1aS80yThNT3iP4T00/edit#gid=978414281"
anno_df <- read_sheet(anno_url, "transcript-regions-mark") %>%
  filter(Experiment=="Mixing")
regs <- split(anno_df$Name_in_gtf_file, anno_df$region_name)


# compute aggregates for egfp/tdtomato for TSO
ae <- altExp(sce_filt,"tso_ontarget_multi")
cd <- colData(ae)

for(i in 1:length(regs))
  cd[[ names(regs)[i] ]] <- colSums(assay(ae,1)[regs[[i]],,drop=FALSE])

colData(altExp(sce_filt,"tso_ontarget_multi")) <- cd
# ---------------

# compute aggregates for egfp/tdtomato for WTA
ae <- altExp(sce_filt,"wta_ontarget_multi")
cd <- colData(ae)

for(i in 1:length(regs))
  cd[[ names(regs)[i] ]] <- colSums(assay(ae,1)[regs[[i]],,drop=FALSE])

colData(altExp(sce_filt,"wta_ontarget_multi")) <- cd
# ---------------

```

### Score regions tso_multi filt dataset 

```{r}

ae <- altExp(sce_filt,"tso_ontarget_multi",withColData=TRUE)

ae$tso_egfp_detection_score <- 0
ae$tso_egfp_detection_score[ae$egfp_tx>0] <- 1
ae$tso_egfp_detection_score[ae$egfp_cds>0] <- 2
ae$tso_egfp_detection_score[ae$egfp_roi>0] <- 3
ae$tso_egfp_detection_score[ae$egfp_roi>0] <- 4

ae$tso_tdtom_detection_score<-0
ae$tso_tdtom_detection_score[ae$tdtom_tx>0] <- 1
ae$tso_tdtom_detection_score[ae$tdtom_cds>0] <- 2
ae$tso_tdtom_detection_score[ae$tdtom_roi1>0] <- 3
ae$tso_tdtom_detection_score[ae$tdtom_roi2>0] <- 3
ae$tso_tdtom_detection_score[ae$tdtom_roi2>0 & ae$tdtom_roi1>0] <- 4

table(ae$tso_egfp_detection_score)
table(ae$tso_tdtom_detection_score)

altExp(sce_filt,"tso_ontarget_multi") <- ae

```

### Score regions wta_multi filt dataset 

```{r}

ae <- altExp(sce_filt,"wta_ontarget_multi",withColData=TRUE)

ae$wta_egfp_detection_score <- 0
ae$wta_egfp_detection_score[ae$egfp_tx>0] <- 1
ae$wta_egfp_detection_score[ae$egfp_cds>0] <- 2
ae$wta_egfp_detection_score[ae$egfp_roi>0] <- 3
ae$wta_egfp_detection_score[ae$egfp_roi>0] <- 4

ae$wta_tdtom_detection_score <- 0
ae$wta_tdtom_detection_score[ae$tdtom_tx>0] <- 1
ae$wta_tdtom_detection_score[ae$tdtom_cds>0] <- 2
ae$wta_tdtom_detection_score[ae$tdtom_roi1>0] <- 3
ae$wta_tdtom_detection_score[ae$tdtom_roi2>0] <- 3
ae$wta_tdtom_detection_score[ae$tdtom_roi2>0 & ae$tdtom_roi1>0] <- 4

table(ae$wta_egfp_detection_score)
table(ae$wta_tdtom_detection_score)

altExp(sce_filt,"wta_ontarget_multi") <- ae


```

### Splitting mouse vs human filt dataset and removing cells with unknown species

```{r}

sce_filt$species <- "unknown"
sce_filt$species[sce_filt$subsets_mouse_gene_percent>50] <- "mouse"
sce_filt$species[sce_filt$subsets_human_gene_percent>50] <- "human"

sce_filt <- sce_filt[, !sce_filt$species== "unknown"]

cd <- colData(sce_filt) %>% as.data.frame %>% 
  mutate(total_mito_percent = subsets_mouse_mito_percent+subsets_human_mito_percent)

table(sce_filt$species,
      altExp(sce_filt,"tso_ontarget_multi",withColData=TRUE)$tso_egfp_detection_score)
table(sce_filt$species,
      altExp(sce_filt,"tso_ontarget_multi",withColData=TRUE)$tso_tdtom_detection_score)

```

### Barnyard plot for tso data --> in this case just care about the cds and rois, not transcript

```{r}

ae_tso <- altExp(sce_filt,"tso_ontarget_multi")
ae_wta <- altExp(sce_filt,"wta_ontarget_multi")

comb <- data.frame(colData(sce_filt)[,c("subsets_mouse_gene_sum","subsets_human_gene_sum",
                                        "sample_id")],
                   colData(ae_tso)[,c("tso_tdtom_detection_score","tso_egfp_detection_score")],
                   colData(ae_wta)[,c("wta_tdtom_detection_score","wta_egfp_detection_score")])

head(comb, 2)

brewer.pal(n=4,name="BuGn")
brewer.pal(n=4,name="Blues")

comb$color<-"purple" # these will be the doublets
comb$color[comb$tso_egfp_detection_score==0 & comb$tso_tdtom_detection_score==0]<-"grey" # no detection of either
comb$color[comb$tso_egfp_detection_score==1 & comb$tso_tdtom_detection_score>0]<-"grey" # no detection of either
comb$color[comb$tso_egfp_detection_score>0 & comb$tso_tdtom_detection_score==1]<-"grey" # no detection of either

comb$color[comb$tso_egfp_detection_score==2 & comb$tso_tdtom_detection_score<=1]<-"#B2E2E2"
comb$color[comb$tso_egfp_detection_score==3 & comb$tso_tdtom_detection_score<=1]<-"#66C2A4"
comb$color[comb$tso_egfp_detection_score==4 & comb$tso_tdtom_detection_score<=1]<-"#238B45"

comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==2]<-"#BDD7E7"
comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==3]<-"#6BAED6"
comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==4]<-"#2171B5"

table(comb$color,comb$sample_id)

ggplot(comb, aes(
  x = subsets_mouse_gene_sum,
  y = subsets_human_gene_sum,
  color = color 
)) +
  geom_point(alpha = 0.5,size=0.5) +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", 
                                           "mixing_rock", "mixing_rockroi"))) +
  theme_bw() +
  xlab("mouse gene sum") +
  ylab("human gene sum") +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 20, face = "bold"),
    strip.text = element_text(size = 15, face = "bold")
  ) +
  scale_color_identity(
    name = "Legend Title",
    guide = "legend" ,
  )+
  labs(title="detection of egfp and tdtomato regions")+
  scale_color_manual(labels=c("all_tdtom_rois","all_egfp_rois","tdtom_roi",
                              "egfp_cds","tdtom_cds","no_detection","doublets"),
                     values=c("#2171B5","#238B45","#6BAED6","#B2E2E2","#BDD7E7","grey","purple"))

```

### Barnyard plot with ROI detection 

```{r}

comb$color<-"purple" # these will be the doublets
comb$color[comb$tso_egfp_detection_score==0 & comb$tso_tdtom_detection_score==0]<-"grey" # no detection of either
comb$color[comb$tso_egfp_detection_score==1 & comb$tso_tdtom_detection_score>0]<-"grey" # no detection of either
comb$color[comb$tso_egfp_detection_score>0 & comb$tso_tdtom_detection_score==1]<-"grey" # no detection of either

comb$color[comb$tso_egfp_detection_score==2 & comb$tso_tdtom_detection_score<=1]<-"grey"
comb$color[comb$tso_egfp_detection_score==3 & comb$tso_tdtom_detection_score<=1]<-"#66C2A4"
comb$color[comb$tso_egfp_detection_score==4 & comb$tso_tdtom_detection_score<=1]<-"#238B45"

comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==2]<-"grey"
comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==3]<-"#6BAED6"
comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==4]<-"#2171B5"

ggplot(comb, aes(
  x = subsets_mouse_gene_sum,
  y = subsets_human_gene_sum,
  color = color 
)) +
  geom_point(alpha = 0.5,size = 0.5) +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", 
                                           "mixing_rock", "mixing_rockroi"))) +
  theme_bw() +
  xlab("mouse gene sum") +
  ylab("human gene sum") +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 20, face = "bold"),
    strip.text = element_text(size = 15, face = "bold")
  ) +
  scale_color_identity(
    name = "Legend Title",
    guide = "legend" ,
  )+
  labs(title="detection of egfp and tdtomato regions")+
  scale_color_manual(labels=c("all_tdtom_rois","all_egfp_rois","tdtom_roi",
                              "no_detection","doublets"), 
                     values=c("#2171B5","#238B45","#6BAED6","grey","purple"))

```

### Barnyard plot with cds detection

```{r}

comb$color<-"purple" # these will be the doublets
comb$color[comb$tso_egfp_detection_score==0 & comb$tso_tdtom_detection_score==0]<-"grey" # no detection of either
comb$color[comb$tso_egfp_detection_score==1 & comb$tso_tdtom_detection_score>0]<-"grey" # no detection of either
comb$color[comb$tso_egfp_detection_score>0 & comb$tso_tdtom_detection_score==1]<-"grey" # no detection of either

comb$color[comb$tso_egfp_detection_score==2 & comb$tso_tdtom_detection_score<=1]<-"#238B45" # rois are part of the cds
comb$color[comb$tso_egfp_detection_score==3 & comb$tso_tdtom_detection_score<=1]<-"#238B45" # rois are part of the cds
comb$color[comb$tso_egfp_detection_score==4 & comb$tso_tdtom_detection_score<=1]<-"#238B45" # rois are part of the cds

comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==2]<-"#2171B5" # rois are part of the cds
comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==3]<-"#2171B5" # rois are part of the cds
comb$color[comb$tso_egfp_detection_score<=1 & comb$tso_tdtom_detection_score==4]<-"#2171B5" # rois are part of the cds

ggplot(comb, aes(
  x = subsets_mouse_gene_sum,
  y = subsets_human_gene_sum,
  color = color 
)) +
  geom_point(alpha = 0.5,size=0.5) +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", "mixing_rock", "mixing_rockroi"))) +
  theme_bw() +
  xlab("mouse gene sum") +
  ylab("human gene sum") +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 20, face = "bold"),
    strip.text = element_text(size = 15, face = "bold")
  ) +
  scale_color_identity(
    name = "Legend Title",
    guide = "legend" ,
  )+
  labs(title="detection of egfp and tdtomato regions")+
  scale_color_manual(labels=c("tdtom_cds","egfp_cds","no_detection","doublets"), values=c("#2171B5","#238B45","grey","purple"))

```

### Removing doublets

```{r}

sce_filt$doublet <- FALSE
sce_filt$doublet[sce_filt$subsets_human_gene_percent > 10 & 
               sce_filt$subsets_mouse_gene_percent > 10] <- TRUE

table(sce_filt$doublet)

# ggplot(colData(sce_filt) %>% as.data.frame,
#        aes(x = subsets_mouse_gene_percent, 
#            y = subsets_human_gene_percent,
#            colour = doublet)) +
#   geom_point() + facet_wrap(~sample_id) + geom_density2d(colour="white")

sce_nd <- sce_filt[,!sce_filt$doublet]

rm(sce_filt)

```

### TSO vs WTA plot --> color by on target tso_ontarget_multi

```{r}

# on target TSO multi --> sum of counts for whole transcript

rds <- rowData(sce_nd)

tso_qc <- perCellQCMetrics(sce_nd, assay.type = "tso_off_and_ontarget_unique",
                           subsets=list(capture=rds$type=="capture",
                                        human_gene=rds$type=="human_gene",
                                        human_mito=rds$type=="human_mito",
                                        mouse_gene=rds$type=="mouse_gene",
                                        mouse_mito=rds$type=="mouse_mito"))

df <- data.frame(tso_total = tso_qc$total,
                 tso_capture_sum = tso_qc$subsets_capture_sum,
                 tso_mito_percent = tso_qc$subsets_mouse_mito_percent+
                   tso_qc$subsets_human_mito_percent,
                 wta_total = sce_nd$total,
                 wta_capture_sum = sce_nd$subsets_capture_sum,
                 wta_mito_percent = sce_nd$subsets_mouse_mito_percent+
                   sce_nd$subsets_human_mito_percent,
                 sample_id = sce_nd$sample_id)


colData(altExp(sce_nd,"tso_ontarget_multi",withColData=TRUE))$tso_capture_sum<-colSums(assay(altExp(sce_filt,"tso_ontarget_multi"),1)[c("5_utr_egfp_wpre","3_utr_egfp_wpre","5_to_roi_egfp","roi_egfp","rock_egfp","capture_sequence_double_egfp_egfp"),]) # excluding regulatory region 
colData(altExp(sce_nd,"wta_ontarget_multi",withColData=TRUE))$wta_capture_sum<-colSums(assay(altExp(sce_filt,"wta_ontarget_multi"),1)[c("5_utr_egfp_wpre","3_utr_egfp_wpre","5_to_roi_egfp","roi_egfp","rock_egfp","capture_sequence_double_egfp_egfp"),])

log_1p_tso_ontarget_multi <- log1p(colData(altExp(sce_nd,"tso_ontarget_multi",withColData=TRUE))$tso_capture_sum)

brewer.pal(n=9,name="BuGn")

ggplot(df, aes(x = tso_total, y = wta_total, color = log_1p_tso_ontarget_multi)) +
  geom_point(alpha=0.3,size=0.5) +
  scale_x_log10() + scale_y_log10() +
  theme_bw() +
  xlab("total TSO UMIs") +
  ylab("total WTA UMIs") +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 20, face = "bold"),
    strip.text = element_text(size = 15, face = "bold"))+
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", "mixing_rock", "mixing_rockroi")))+
  labs(title="total WTA UMIs vs total TSO UMIs")+
  scale_color_viridis()

```

### UMIs TSO vs WTA capture, colored by tso_ontarget_multi ordered categorical

```{r}

# adding the score to sces

df$tso_egfp_detection_score<-altExp(sce_nd,"tso_ontarget_multi",withColData=TRUE)$tso_egfp_detection_score
df$tso_tdtom_detection_score<-altExp(sce_nd,"tso_ontarget_multi",withColData=TRUE)$tso_tdtom_detection_score

df$color<-"purple" # these will be the doublets
df$color[df$tso_egfp_detection_score==0 & df$tso_tdtom_detection_score==0]<-"grey" # no detection of either
df$color[df$tso_egfp_detection_score==1 & df$tso_tdtom_detection_score>0]<-"grey" # no detection of either
df$color[df$tso_egfp_detection_score>0 & df$tso_tdtom_detection_score==1]<-"grey" # no detection of either

df$color[df$tso_egfp_detection_score==2 & df$tso_tdtom_detection_score<=1]<-"#B2E2E2"
df$color[df$tso_egfp_detection_score==3 & df$tso_tdtom_detection_score<=1]<-"#66C2A4"
df$color[df$tso_egfp_detection_score==4 & df$tso_tdtom_detection_score<=1]<-"#238B45"

df$color[df$tso_egfp_detection_score<=1 & df$tso_tdtom_detection_score==2]<-"#BDD7E7"
df$color[df$tso_egfp_detection_score<=1 & df$tso_tdtom_detection_score==3]<-"#6BAED6"
df$color[df$tso_egfp_detection_score<=1 & df$tso_tdtom_detection_score==4]<-"#2171B5"

ggplot(df, aes(x = tso_capture_sum, y = wta_capture_sum, color=color))+
  geom_jitter(height = .2,size=0.5,alpha=0.5)+
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", "mixing_rock", "mixing_rockroi"))) +
  theme_bw()+
  xlab("on target TSO UMIs")+
  ylab("on target WTA UMIs")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  labs(title="detection of egfp and tdtomato regions")+
  scale_color_manual(labels=c("all_tdtom_rois","all_egfp_rois","tdtom_roi",
                              "egfp_cds","tdtom_cds","no_detection","doublets"),
                     values=c("#2171B5","#238B45","#6BAED6","#B2E2E2","#BDD7E7","grey","purple"))

```

### Checking egfp ROIs

```{r}

table(df$tso_egfp_detection_score,sce_nd$sample_id,sce_nd$species)

ggplot(df, aes(x = tso_capture_sum, y = tso_egfp_detection_score,color=wta_total))+
  geom_jitter(height = .2,size=0.5,alpha=0.5)+
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", 
                                           "mixing_rock", "mixing_rockroi"))) +
  theme_bw()+
  xlab("on target TSO UMIs")+
  ylab("tso_egfp_detection_score")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  scale_color_viridis()

```

### Checking tdtom ROIs

```{r}

table(df$tso_tdtom_detection_score,sce_nd$sample_id,sce_nd$species)

ggplot(df, aes(x = tso_capture_sum, y = tso_tdtom_detection_score,color=wta_total))+
  geom_jitter(height = .2,size=0.5,alpha=0.5)+
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", 
                                           "mixing_rock", "mixing_rockroi"))) +
  theme_bw()+
  xlab("on target TSO UMIs")+
  ylab("tso_tdtom_detection_score")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  scale_color_viridis()

```

### On target UMIs TSO vs WTA capture, colored by raw total reads

```{r}

ggplot(df, aes(x = tso_capture_sum, y = wta_capture_sum, color=wta_total))+
  geom_jitter(height = .2,alpha=0.3,size=0.5)+
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", 
                                           "mixing_rock", "mixing_rockroi"))) +
  theme_bw()+
  xlab("on target TSO UMIs")+
  ylab("on target WTA UMIs")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  labs(title="on target WTA and TSO UMIs")+
  scale_color_viridis()
  
```

### On target UMIs TSO vs WTA, colored by CDS

```{r}

df$color<-"purple" # these will be the doublets
df$color[df$tso_egfp_detection_score==0 & df$tso_tdtom_detection_score==0]<-"grey" # no detection of either
df$color[df$tso_egfp_detection_score==1 & df$tso_tdtom_detection_score>0]<-"grey" # no detection of either
df$color[df$tso_egfp_detection_score>0 & df$tso_tdtom_detection_score==1]<-"grey" # no detection of either

df$color[df$tso_egfp_detection_score==2 & df$tso_tdtom_detection_score<=1]<-"#238B45"
df$color[df$tso_egfp_detection_score==3 & df$tso_tdtom_detection_score<=1]<-"#238B45"
df$color[df$tso_egfp_detection_score==4 & df$tso_tdtom_detection_score<=1]<-"#238B45"

df$color[df$tso_egfp_detection_score<=1 & df$tso_tdtom_detection_score==2]<-"#2171B5"
df$color[df$tso_egfp_detection_score<=1 & df$tso_tdtom_detection_score==3]<-"#2171B5"
df$color[df$tso_egfp_detection_score<=1 & df$tso_tdtom_detection_score==4]<-"#2171B5"

ggplot(df, aes(x = tso_capture_sum, y = wta_capture_sum, color=color))+
  geom_jitter(height = .2,alpha=0.3,size=0.5)+
  facet_wrap(~factor(sample_id, levels = c("mixing_unmod", "mixing_unmod_roi", "mixing_rock", "mixing_rockroi"))) +
  theme_bw()+
  xlab("on target TSO UMIs")+
  ylab("on target WTA UMIs")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  labs(title="on target WTA and TSO UMIs")+
  scale_color_manual(labels=c("tdtom_cds","egfp_cds","no_detection","doublets"), values=c("#2171B5","#238B45","grey","purple"))

```

### False positive vs true positive

```{r}

exprs_nd <- assay(altExp(sce_nd,"tso_ontarget_multi",withColData=TRUE),1)

sort(rowSums(exprs_nd, na.rm = TRUE), decreasing = TRUE) -> sorted

df <- data.frame(mouse = (sce_nd$species == "mouse")+0,
                 human = (sce_nd$species == "human")+0,
                 sample = sce_nd$sample_id,
                 as.matrix(t(exprs_nd[names(sorted),])))

dfm <- melt(df, id.vars = c("human","mouse","sample"))

human_targets <- grep("tdtomato", names(sorted), value = TRUE)

p <- ggplot(dfm %>% filter(variable %in% c("roi_1_tdtomato", "roi_2_tdtomato","rock_tdtomato")),
                      aes(m = value, d = human, colour=sample, group=sample)) + 
  geom_roc(size = 1.2, alpha = .8,
           position = position_dodge2(width=.001)) +
  facet_grid(~variable) +
  xlim(0, .018) +
  ggtitle("Human targets") +
  theme(legend.position="bottom") +
  theme_bw()+
  xlab("false positive fraction")+
  ylab("true positive fraction")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))
  
mouse_targets <- grep("egfp", names(sorted), value = TRUE)

q <- ggplot(dfm %>% filter(variable %in% c("rock_egfp","roi_egfp")),
                      aes(m = value, d = mouse, colour=sample, group=sample)) + 
  geom_roc(size = 1.2, alpha = .8,
           position = position_dodge2(width=.001)) +
  facet_grid(~variable) +
  xlim(0, .018) +
  ggtitle("Mouse targets") +
  theme(legend.position="none") +
  theme_bw()+
  xlab("false positive fraction")+
  ylab("true positive fraction")+
  theme(axis.text = element_text(size = 15),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))
  
cowplot::plot_grid(p,q, nrow=2)

```


### Fuzznuc ROI

```{r}

# Getting error for Matrix package

```

### Barplot with % positive cells (for now just tso_ontarget_multi, so now just for cds)

```{r}

alt<-altExp(sce_nd,withColData=TRUE)

mouse<-alt[,alt$species=="mouse"]
human<-alt[,alt$species=="human"]

positive_cells<-data.frame(matrix(nrow = 4, ncol = 2)) 
colnames(positive_cells)<-c("percent_egfp_pos_mouse","percent_tdtom_pos_human")
rownames(positive_cells)<-unique(alt$sample_id)

for (i in unique(mouse$sample_id)){
  positive_cells[i,1]<-length(which(mouse[,mouse$sample_id==i]$egfp_cds>0))/length(mouse[,mouse$sample_id==i]$egfp_cds)*100
}

for (i in unique(human$sample_id)){
  positive_cells[i,2]<-length(which(human[,human$sample_id==i]$tdtom_cds>0))/length(human[,human$sample_id==i]$tdtom_cds)*100
}

plottable<-melt(positive_cells)
plottable$sample<-rep(rownames(positive_cells),2)

ggplot(data=plottable,aes(y=value,x=variable,fill=factor(sample,levels=c("mixing_unmod","mixing_unmod_roi","mixing_rock","mixing_rockroi")))) +
  geom_bar(stat="identity",width = 0.8, position = position_dodge(width = 0.9))+
  xlab(" ")+
  ylab("percent positive cells")+
  theme_bw()+
  theme(axis.text = element_text(size = 10),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  scale_fill_discrete(name = "Samples")+
  geom_text(aes(label=format(round(value,2),nsmall=2)), position=position_dodge(width=0.9), vjust=-0.5)

```

### Scatter plot with % positive cells (for now just tso_ontarget_multi, so now just for cds)

```{r}

ggplot(data=plottable,aes(y=value,x=variable,color=factor(sample,levels=c("mixing_unmod","mixing_unmod_roi","mixing_rock","mixing_rockroi")))) +
  geom_point()+
  xlab(" ")+
  ylab("percent positive cells")+
  theme_bw()+
  labs(color='Samples') +
  theme(axis.text = element_text(size = 10),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  geom_text(aes(label=format(round(value,2),nsmall=2)), position=position_dodge(width=0.5), vjust=-0.5,show.legend=FALSE,size=4)
  
```

### Scatter plot with % positive cells with different x axis (for now just tso_ontarget_multi, so now just for cds)

```{r}

ggplot(data=plottable,aes(y=value,x=factor(sample,levels=c("mixing_unmod","mixing_unmod_roi","mixing_rock","mixing_rockroi")),color=variable)) +
  geom_point()+
  xlab(" ")+
  ylab("percent positive cells")+
  theme_bw()+
  labs(color='Samples') +
  theme(axis.text = element_text(size = 10),axis.title = element_text(size = 20, face = "bold"),strip.text = element_text(size = 15, face = "bold"))+
  geom_text(aes(label=format(round(value,2),nsmall=2)), position=position_dodge(width=0.5), vjust=-0.5,show.legend=FALSE,size=4)+
  scale_color_manual(labels=c("percent_egfp_pos_mouse","percent_tdtom_pos_human"), values=c("#2171B5","#238B45"))

```

### Library size per cell WTA vs TSO (previous plots from Izaskun)

```{r}

rds <- rowData(sce_nd)

tso_qc <- perCellQCMetrics(sce_nd, assay.type = "tso_off_and_ontarget_unique",
                           subsets=list(capture=rds$type=="capture",
                                        human_gene=rds$type=="human_gene",
                                        human_mito=rds$type=="human_mito",
                                        mouse_gene=rds$type=="mouse_gene",
                                        mouse_mito=rds$type=="mouse_mito"))

df <- data.frame(tso_total = tso_qc$total,
                 tso_capture_sum = tso_qc$subsets_capture_sum,
                 tso_mito_percent = tso_qc$subsets_mouse_mito_percent+
                   tso_qc$subsets_human_mito_percent,
                 wta_total = sce_nd$total,
                 wta_capture_sum = sce_nd$subsets_capture_sum,
                 wta_mito_percent = sce_nd$subsets_mouse_mito_percent+
                   sce_nd$subsets_human_mito_percent,
                 sample_id = sce_nd$sample_id)


df$tso_capture_sum<-colSums(assay(altExp(sce_filt,"tso_ontarget_multi"),1)[c("5_utr_egfp_wpre","3_utr_egfp_wpre","5_to_roi_egfp","roi_egfp","rock_egfp","capture_sequence_double_egfp_egfp"),])
df$wta_capture_sum<-colSums(assay(altExp(sce_filt,"wta_ontarget_multi"),1)[c("5_utr_egfp_wpre","3_utr_egfp_wpre","5_to_roi_egfp","roi_egfp","rock_egfp","capture_sequence_double_egfp_egfp"),])

samples<-c("mixing_unmod","mixing_unmod_roi","mixing_rock","mixing_rockroi")

p <- list()
for (i in samples){
     p[[i]]<-ggplot(df[df$sample_id==i,], aes(x = wta_capture_sum,
                    y = tso_capture_sum,
                    color = wta_total)) +
      geom_point(alpha = 0.1) +
      geom_smooth() +
      scale_x_continuous(trans='log1p') +
      scale_y_continuous(trans='log1p') +
      theme_bw() +
      labs(title=i)+
      theme(aspect.ratio = 1, legend.position = "bottom") +
      scale_color_gradient(trans = "log1p") +
      scale_color_viridis() +
      xlab('on-target WTA # UMIs') +
      ylab('on-target TSO # UMIs')
     
      p[[i]]<-ggExtra::ggMarginal(p[[i]],type = "density")
}

do.call(grid.arrange,p)

```

### Rendering file

```{r}

render("/Users/gimoro/figure_making_paper/rock_roi_paper/00_mixing_experiment/qmd/00_mixing_experiment_c-figures.qmd","html_document")

```

### Print sessionInfo()

```{r}

sessionInfo()

```


