---
title: "00_mixing_experiment_d-coverage_tracks.qmd"
author: "Mark Robinson, Izaskun Mallona & Giulia Moro"
format: 
  html:
    toc: true
    embed-resources: true
    keep_md: true
---


### TO DO 

```{r}

# Change objects to subset with two peaks 

```

### Setting up knitr 

```{r}

library(knitr)

knitr::opts_chunk$set(fig.width = 14,
               fig.height = 8,
               fig.align = "center",
               include = TRUE,
               dev = c("png","svg"),
               cache.lazy = FALSE,
               cache=TRUE,
               warning = TRUE,
               message = TRUE)

```

### Load packages, load data 

```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(GenomicRanges)
library(GenomicAlignments)
library(rtracklayer)
library(Gviz)
library(plyranges)
library(seqinr)
library(Rsamtools)

```

### Reading .bam files

```{r}

getwd()

bw<-dir(here("00_mixing_experiment"), "coverage.*bw", 
            recursive = TRUE, full.names = TRUE)

names(bw) <- gsub("_coverage.bw", "", basename(bw))
names(bw) <- gsub("mixing_", "",names(bw))

datadir <- dirname(bw)[1] # folder where the .bw files are 

```

### Importing .gtf file and subsetting for egfp and tdtomato

```{r}

options(ucscChromosomeNames=FALSE)

gtf<-dir(here("00_mixing_experiment"), ".*gtf", 
            recursive = TRUE, full.names = TRUE)

combined_gtf<-import(gtf)
gtf_egfp<-combined_gtf[which(seqnames(combined_gtf)=="egfp"),]
gtf_tdtom<-combined_gtf[which(seqnames(combined_gtf)=="tdtomato"),]

```

### Importing excel sheet

```{r}

# read annotation info from GS
anno_url <- "https://docs.google.com/spreadsheets/d/1cvldF_VFA7FYuChoR9a4SIpYcZ1aS80yThNT3iP4T00/edit#gid=978414281"
anno_df <- read_sheet(anno_url, "transcript-regions-mark") %>%
  filter(Experiment=="Mixing")
regs <- split(anno_df$Name_in_gtf_file, anno_df$region_name)

```

### Plots for all .bw files for egfp 

```{r}

chr<-c("egfp")

coverage_egfp <- lapply(bw, function(x) {
  import.bw(x, as="GRanges")
  })

for (i in 1:length(coverage_egfp)){
  coverage_egfp[[i]]<-DataTrack(coverage_egfp[[i]],
                                chromosome=chr,
                                name=names(coverage_egfp[i]))
}

gm_egfp <- GeneRegionTrack(gtf_egfp,
                           chromosome=chr,
                           name=chr,
                           geneSymbol=TRUE,
                           showId=TRUE,
                           transcriptAnnotation="gene",
                           fill="cadetblue2")

gm_egfp@range$gene<-gtf_egfp$gene_id
wanted_regions<-regs$egfp_tx
gm_egfp@range<-gm_egfp@range[which(gm_egfp@range$gene %in% wanted_regions),]

displayPars(gm_egfp)<-list(col="black",
                           cex.group=0.8,
                           ol.line="black",
                           fontcolor.group="black",
                           add53=TRUE,)


for (i in names(coverage_egfp)){
  displayPars(coverage_egfp[[i]])<-list(col.histogram="black",fill.histogram=TRUE)
}

plotTracks(c(coverage_egfp$unmod_wta,
             coverage_egfp$unmod_tso,
             coverage_egfp$unmod_roi_wta,
             coverage_egfp$unmod_roi_tso,
             coverage_egfp$rock_wta,
             coverage_egfp$rock_tso,
             coverage_egfp$rockroi_wta,
             coverage_egfp$rockroi_tso,
             gm_egfp),
           type="histogram",
           from=start(gm_egfp@range)[1]-400,
           end=end,
           cex.axis = 0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5)

```

### Same for tdtomato for all .bw files

```{r}

chr<-c("tdtomato")

coverage_tdtomato <- lapply(bw, function(x) {
  import.bw(x, as="GRanges")
  })

for (i in 1:length(coverage_tdtomato)){
  coverage_tdtomato[[i]]<-DataTrack(coverage_tdtomato[[i]],
                                    chromosome=chr,
                                    name=names(coverage_tdtomato[i]))
}

gm_tdtomato <- GeneRegionTrack(gtf_tdtom,
                               chromosome=chr,
                               name=chr,
                               geneSymbol=TRUE,
                               showId=TRUE,
                               transcriptAnnotation="gene",
                               fill="cadetblue2")

gm_tdtomato@range$gene<-gtf_tdtom$gene_id

wanted_regions<-regs$tdtom_tx
gm_tdtomato@range<-gm_tdtomato@range[which(gm_tdtomato@range$gene %in% wanted_regions),]

displayPars(gm_tdtomato)<-list(col="black",cex.group=0.8,col.line="black",fontcolor.group="black")

for (i in names(coverage_tdtomato)){
  displayPars(coverage_tdtomato[[i]])<-list(col.histogram="black",fill.histogram=TRUE)
}

plotTracks(c(coverage_tdtomato$unmod_wta,
             coverage_tdtomato$unmod_tso,
             coverage_tdtomato$unmod_roi_wta,
             coverage_tdtomato$unmod_roi_tso,
             coverage_tdtomato$rock_wta,
             coverage_tdtomato$rock_tso,
             coverage_tdtomato$rockroi_wta,
             coverage_tdtomato$rockroi_tso,
             gm_tdtomato),
           type="histogram",
           from=start(gm_tdtomato@range)[1]-400,
           end=end,
           cex.axis = 0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5)

```

### Putting wta and tso tracks together for egfp

```{r}

ylims_unmod_egfp <- extendrange(range(c(values(coverage_egfp$unmod_wta), values(coverage_egfp$unmod_tso))))
displayPars(coverage_egfp$unmod_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$unmod_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_unmod_roi_egfp <- extendrange(range(c(values(coverage_egfp$unmod_roi_wta), values(coverage_egfp$unmod_roi_tso))))
displayPars(coverage_egfp$unmod_roi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$unmod_roi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rock_egfp <- extendrange(range(c(values(coverage_egfp$rock_wta), values(coverage_egfp$rock_tso))))
displayPars(coverage_egfp$rock_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$rock_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rockroi_egfp <- extendrange(range(c(values(coverage_egfp$rockroi_wta), values(coverage_egfp$rockroi_tso))))
displayPars(coverage_egfp$rockroi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$rockroi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

for (i in 1:length(coverage_egfp)){
  coverage_egfp[[i]]@name<-gsub(c("_wta","_tso"),"",coverage_egfp[[i]]@name)
}

unmod_egfp<-OverlayTrack(trackList=list(coverage_egfp$unmod_wta,coverage_egfp$unmod_tso),legend=TRUE)
unmod_roi_egfp<-OverlayTrack(trackList=list(coverage_egfp$unmod_roi_wta,coverage_egfp$unmod_roi_tso),legend=TRUE)
rock_egfp<-OverlayTrack(trackList=list(coverage_egfp$rock_wta,coverage_egfp$rock_tso),legend=TRUE,name=)
rockroi_egfp<-OverlayTrack(trackList=list(coverage_egfp$rockroi_wta,coverage_egfp$rockroi_tso),legend=TRUE)

plotTracks(c(unmod_egfp,
             unmod_roi_egfp,
             rock_egfp,
             rockroi_egfp,
             gm_egfp),
           type="histogram",
           from=start(gm_egfp@range)[1]-400,
           end=end,
           cex.axis = 0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5,
           ylim=c(0,max(ylims_rockroi_egfp,
                        ylims_rock_egfp,
                        ylims_unmod_roi_egfp,
                        ylims_unmod_egfp)))

```

### Putting wta and tso tracks together for tdtomato

```{r}

ylims_unmod_tdtomato <- extendrange(range(c(values(coverage_tdtomato$unmod_wta), values(coverage_tdtomato$unmod_tso))))
displayPars(coverage_tdtomato$unmod_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$unmod_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_unmod_roi_tdtomato <- extendrange(range(c(values(coverage_tdtomato$unmod_roi_wta), values(coverage_tdtomato$unmod_roi_tso))))
displayPars(coverage_tdtomato$unmod_roi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$unmod_roi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rock_tdtomato <- extendrange(range(c(values(coverage_tdtomato$rock_wta), values(coverage_tdtomato$rock_tso))))
displayPars(coverage_tdtomato$rock_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$rock_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rockroi_tdtomato <- extendrange(range(c(values(coverage_tdtomato$rockroi_wta), values(coverage_tdtomato$rockroi_tso))))
displayPars(coverage_tdtomato$rockroi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$rockroi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

for (i in 1:length(coverage_egfp)){
  coverage_tdtomato[[i]]@name<-gsub(c("_wta","_tso"),"",coverage_tdtomato[[i]]@name)
}

unmod_tdtomato <-OverlayTrack(trackList=list(coverage_tdtomato$unmod_wta,coverage_tdtomato$unmod_tso),legend=TRUE)
unmod_roi_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$unmod_roi_wta,coverage_tdtomato$unmod_roi_tso),legend=TRUE)
rock_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$rock_wta,coverage_tdtomato$rock_tso),legend=TRUE)
rockroi_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$rockroi_wta,coverage_tdtomato$rockroi_tso),legend=TRUE)

plotTracks(c(unmod_tdtomato,
             unmod_roi_tdtomato,
             rock_tdtomato,
             rockroi_tdtomato,
             gm_tdtomato),
           type="histogram",
           showSampleNames = TRUE,
           from=start(gm_egfp@range)[1]-400,
           end=end,
           cex.axis = 0.5,
           cex.title= 0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5,
           ylim=c(0,max(ylims_rockroi_tdtomato,
                        ylims_rock_tdtomato,
                        ylims_unmod_roi_tdtomato,
                        ylims_unmod_tdtomato)))

```

### Reading .bam files

```{r}

getwd()

bam<-dir(here("00_mixing_experiment"), "*.bam", 
            recursive = TRUE, full.names = TRUE)

names(bam) <- gsub(".bam", "", basename(bam))
names(bam) <- gsub("_subset", "",names(bam))

datadir <- dirname(bam)[1] # folder where the .bam files are 

for (i in 1:length(bam)){
  indexBam(bam[[i]])
}

```

### Same for .bam files for eGFP

```{r}

options(ucscChromosomeNames=FALSE)

chr<-c("egfp")

z_egfp <- GRanges(chr, IRanges(start(gm_egfp@range)[1]-400,max(end(gm_egfp@range))))
param <- ScanBamParam(which=z_egfp)

gals_egfp <- lapply(bam, function(x) {
  readGAlignments(x, use.names = TRUE, param=param)
  })

covs_egfp<-lapply(gals_egfp,coverage)
covs_egfp<-lapply(covs_egfp,function(x) as(x,"GRanges"))

for (i in 1:length(covs_egfp)){
  covs_egfp[[i]]<-DataTrack(covs_egfp[[i]],
                            chromosome=chr,
                            name=names(covs_egfp)[[i]])
}

displayPars(gm_egfp)<-list(col="black",cex.group=0.8,col.line="black",fontcolor.group="black")

for (i in names(covs_egfp)){
  displayPars(covs_egfp[[i]])<-list(col.histogram="black",fill.histogram=TRUE)
}

unmod_egfp_bam<-OverlayTrack(trackList=list(covs_egfp$unmod_wta,covs_egfp$unmod_tso),legend=TRUE)
unmod_roi_egfp_bam<-OverlayTrack(trackList=list(covs_egfp$unmod_roi_wta,covs_egfp$unmod_roi_tso),legend=TRUE)
rock_egfp_bam<-OverlayTrack(trackList=list(covs_egfp$rock_wta,covs_egfp$rock_tso),legend=TRUE)
rockroi_egfp_bam<-OverlayTrack(trackList=list(covs_egfp$rockroi_wta,covs_egfp$rockroi_tso),legend=TRUE)

plotTracks(c(unmod_egfp_bam,
             unmod_roi_egfp_bam,
             rock_egfp_bam,
             rockroi_egfp_bam,
             gm_egfp),
           type="histogram",
           from=start(gm_egfp@range)[1]-400,
           end=end,
           cex.axis = 0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5)

```

### Same for .bam files for tdtomato

```{r}

options(ucscChromosomeNames=FALSE)

chr<-c("tdtomato")

z_tdtomato <- GRanges(chr, IRanges(start(gm_tdtomato@range)[1]-400,max(end(gm_tdtomato@range))))
param <- ScanBamParam(which=z_tdtomato)

gals_tdtomato <- lapply(bam, function(x) {
  readGAlignments(x, use.names = TRUE, param=param)
  })

covs_tdtomato<-lapply(gals_tdtomato,coverage)
covs_tdtomato<-lapply(covs_tdtomato,function(x) as(x,"GRanges"))

for (i in 1:length(covs_tdtomato)){
  covs_tdtomato[[i]]<-DataTrack(covs_tdtomato[[i]],
                                chromosome=chr,
                                name=names(covs_tdtomato)[[i]])
}

displayPars(gm_tdtomato)<-list(col="black",cex.group=0.8,col.line="black",fontcolor.group="black")

for (i in names(covs_tdtomato)){
  displayPars(covs_tdtomato[[i]])<-list(col.histogram="black",fill.histogram=TRUE)
}

unmod_tdtomato_bam<-OverlayTrack(trackList=list(covs_tdtomato$unmod_wta,covs_tdtomato$unmod_tso),legend=TRUE)
unmod_roi_tdtomato_bam<-OverlayTrack(trackList=list(covs_tdtomato$unmod_roi_wta,covs_tdtomato$unmod_roi_tso),legend=TRUE)
rock_tdtomato_bam<-OverlayTrack(trackList=list(covs_tdtomato$rock_wta,covs_tdtomato$rock_tso),legend=TRUE)
rockroi_tdtomato_bam<-OverlayTrack(trackList=list(covs_tdtomato$rockroi_wta,covs_tdtomato$rockroi_tso),legend=TRUE)

plotTracks(c(unmod_tdtomato_bam,
             unmod_roi_tdtomato_bam,
             rock_tdtomato_bam,
             rockroi_tdtomato_bam),
           type="histogram",
           from=start(gm_tdtomato@range)[1]-400,
           end=end,
           cex.axis = 0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5)

```

### Putting the plots together for egfp

```{r}

ylims_unmod_egfp_bam <- extendrange(range(c(values(covs_egfp$unmod_wta), values(covs_egfp$unmod_tso))))
ylims_unmod_roi_egfp_bam <- extendrange(range(c(values(covs_egfp$unmod_roi_wta), values(covs_egfp$unmod_roi_tso))))
ylims_rock_egfp_bam <- extendrange(range(c(values(covs_egfp$rock_wta), values(covs_egfp$rock_tso))))
ylims_rockroi_egfp_bam <- extendrange(range(c(values(covs_egfp$rockroi_wta), values(covs_egfp$rockroi_tso))))

plotTracks(c(unmod_egfp_bam,
             unmod_egfp,
             unmod_roi_egfp_bam,
             unmod_roi_egfp,
             rock_egfp_bam,
             rock_egfp,
             rockroi_egfp_bam,
             rockroi_egfp,
             gm_egfp),
           type="histogram",
           from=start(gm_egfp@range)[1]-400,
           end=end,
           cex.axis =0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5,
           ylim=c(0,max(ylims_rockroi_egfp,
                        ylims_rock_egfp,
                        ylims_unmod_roi_egfp,
                        ylims_unmod_egfp,
                        ylims_unmod_egfp_bam,
                        ylims_unmod_roi_egfp_bam,
                        ylims_rock_egfp_bam,
                        ylims_rockroi_egfp_bam)))

```

### Putting the plots together for tdtomato

```{r}

ylims_unmod_tdtomato_bam <- extendrange(range(c(values(covs_tdtomato$unmod_wta), values(covs_tdtomato$unmod_tso))))
ylims_unmod_roi_tdtomato_bam <- extendrange(range(c(values(covs_tdtomato$unmod_roi_wta), values(covs_tdtomato$unmod_roi_tso))))
ylims_rock_tdtomato_bam <- extendrange(range(c(values(covs_tdtomato$rock_wta), values(covs_tdtomato$rock_tso))))
ylims_rockroi_tdtomato_bam <- extendrange(range(c(values(covs_tdtomato$rockroi_wta), values(covs_tdtomato$rockroi_tso))))

plotTracks(c(unmod_tdtomato_bam,
             unmod_tdtomato,
             unmod_roi_tdtomato_bam,
             unmod_roi_tdtomato,
             rock_tdtomato_bam,
             rock_tdtomato,
             rockroi_tdtomato_bam,
             rockroi_tdtomato,
             gm_tdtomato),
           type="histogram",
           from=start(gm_tdtomato@range)[1]-400,
           end=end,
           cex.axis =0.5,
           cex.title=0.8,
           background.title="cadetblue2",
           col.axis="black",
           col.title="black",
           col.border.title="black",
           showSampleNames = TRUE,
           lwd.title=0.5,
           ylim=c(0,max(ylims_rockroi_tdtomato,
                        ylims_rock_tdtomato,
                        ylims_unmod_roi_tdtomato,
                        ylims_unmod_tdtomato,
                        ylims_unmod_tdtomato_bam,
                        ylims_unmod_roi_tdtomato_bam,
                        ylims_rock_tdtomato_bam,
                        ylims_rockroi_tdtomato_bam)))

```

### Print sessionInfo()

```{r}

sessionInfo()

```







