---
title: "00_mixing_experiment_d-coverage_tracks.qmd"
author: "Mark Robinson, Izaskun Mallona & Giulia Moro"
format: 
  html:
    toc: true
    embed-resources: true
    keep_md: true
---


### TO DO 

```{r}

# Make same plots for raw .bam files

```

### Setting up knitr 

```{r}

library(knitr)

knitr::opts_chunk$set(fig.width = 14,
               fig.height = 8,
               fig.align = "center",
               include = TRUE,
               dev = c("png","svg"),
               cache.lazy = FALSE,
               cache=TRUE,
               warning = TRUE,
               message = TRUE)

```

### Load packages, load data 

```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(GenomicRanges)
library(GenomicAlignments)
library(rtracklayer)
library(Gviz)
library(plyranges)

```

### Reading .bw files

```{r}

getwd()

bw<-dir(here("00_mixing_experiment"), "coverage.*bw", 
            recursive = TRUE, full.names = TRUE)

names(bw) <- gsub("_coverage.bw", "", basename(bw))
names(bw) <- gsub("mixing_", "",names(bw))

datadir <- dirname(bw)[1] # folder where the .bw files are 

```

### Importing .gtf file and subsetting for egfp and tdtomato

```{r}

options(ucscChromosomeNames=FALSE)

gtf<-dir(here("00_mixing_experiment"), ".*gtf", 
            recursive = TRUE, full.names = TRUE)

combined_gtf<-import(gtf)
gtf_egfp<-combined_gtf[which(seqnames(combined_gtf)=="egfp"),]
gtf_tdtom<-combined_gtf[which(seqnames(combined_gtf)=="tdtomato"),]

```

### Test for one file for egfp and no regulatory region

```{r}

# chr<-c("egfp")
# 
# test<-import.bw(bw[[8]],as="GRanges")
# egfp<-DataTrack(test,chromosome=chr,name="mixing_rockroi")
# 
# gm_egfp <- GeneRegionTrack(gtf_egfp, chromosome=chr,name=chr,geneSymbol=TRUE,showId=TRUE,transcriptAnnotation="gene",fill="cadetblue2")
# gm_egfp@range$gene<-gtf_egfp$gene_id
# wanted_regions<-c("5_utr_egfp_wpre","3_utr_egfp_wpre","5_to_roi_egfp","roi_egfp","rock_egfp","capture_sequence_double_egfp_egfp") # removing egfp regulatory region
# gm_egfp@range<-gm_egfp@range[which(gm_egfp@range$gene %in% wanted_regions),]
# 
# 
# plotTracks(c(egfp,gm_egfp),type="histogram",showSampleNames = TRUE,from=start(gm_egfp@range)[1]-400,end=end,cex.axis = 0.5,background.title="black") # adding 400 bp so that have space for writing for exon
# 
# plotTracks(c(egfp,gm_egfp),type="histogram",showSampleNames = TRUE,background.title = "black") # in this case just showing transcript, no regulatory region 

## visualization --> from Erich chose the first type of plot, with -44 from the start

```

### Plots for all .bw files for egfp

```{r}

chr<-c("egfp")

coverage_egfp <- lapply(bw, function(x) {
  import.bw(x, as="GRanges")
  })

for (i in 1:length(coverage_egfp)){
  coverage_egfp[[i]]<-DataTrack(coverage_egfp[[i]],chromosome=chr,name=names(coverage_egfp[i]))
}

gm_egfp <- GeneRegionTrack(gtf_egfp, chromosome=chr,name=chr,geneSymbol=TRUE,showId=TRUE,transcriptAnnotation="gene",fill="cadetblue2")
gm_egfp@range$gene<-gtf_egfp$gene_id
wanted_regions<-c("5_utr_egfp_wpre","3_utr_egfp_wpre","5_to_roi_egfp","roi_egfp","rock_egfp","capture_sequence_double_egfp_egfp") # removing egfp regulatory region
gm_egfp@range<-gm_egfp@range[which(gm_egfp@range$gene %in% wanted_regions),]

displayPars(gm_egfp)<-list(col="black",cex.group=0.8,col.line="black",fontcolor.group="black")

names(coverage_egfp)

for (i in names(coverage_egfp)){
  print(i)
  displayPars(coverage_egfp[[i]])<-list(col.histogram="black",fill.histogram=TRUE)
}

plotTracks(c(coverage_egfp$unmod_wta,coverage_egfp$unmod_tso,coverage_egfp$unmod_roi_wta,coverage_egfp$unmod_roi_tso,coverage_egfp$rock_wta,coverage_egfp$rock_tso,coverage_egfp$rockroi_wta,coverage_egfp$rockroi_tso,gm_egfp),type="histogram",showSampleNames = TRUE,from=start(gm_egfp@range)[1]-400,end=end,cex.axis = 0.5,cex.title=0.8,background.title="cadetblue2",col.axis="black",col.title="black",col.border.title="black",showSampleNames = TRUE,lwd.title=0.5)

```
### Same for tdtomato for all .bw files

```{r}
chr<-c("tdtomato")

coverage_tdtomato <- lapply(bw, function(x) {
  import.bw(x, as="GRanges")
  })

for (i in 1:length(coverage_tdtomato)){
  coverage_tdtomato[[i]]<-DataTrack(coverage_tdtomato[[i]],chromosome=chr,name=names(coverage_tdtomato[i]))
}

gm_tdtomato <- GeneRegionTrack(gtf_tdtom, chromosome=chr,name=chr,geneSymbol=TRUE,showId=TRUE,transcriptAnnotation="gene",fill="cadetblue2")
gm_tdtomato@range$gene<-gtf_tdtom$gene_id

wanted_regions<-c("5_utr_tdtomato_wpre","3_utr_tdtomato_wpre","5_to_roi_tdtomato","roi_1_tdtomato","roi_2_tdtomato","roi_3_tdtomato","roi_4_tdtomato","between_roi_1","between_roi_2","between_roi_3","rock_tdtomato","capture_sequence_double_egfp_tdtom") # removing tdtomato regulatory region
gm_tdtomato@range<-gm_tdtomato@range[which(gm_tdtomato@range$gene %in% wanted_regions),]

displayPars(gm_tdtomato)<-list(col="black",cex.group=0.8,col.line="black",fontcolor.group="black")

names(coverage_tdtomato)

for (i in names(coverage_tdtomato)){
  print(i)
  displayPars(coverage_tdtomato[[i]])<-list(col.histogram="black",fill.histogram=TRUE)
}

plotTracks(c(coverage_tdtomato$unmod_wta,coverage_tdtomato$unmod_tso,coverage_tdtomato$unmod_roi_wta,coverage_tdtomato$unmod_roi_tso,coverage_tdtomato$rock_wta,coverage_tdtomato$rock_tso,coverage_tdtomato$rockroi_wta,coverage_tdtomato$rockroi_tso,gm_tdtomato),type="histogram",showSampleNames = TRUE,from=start(gm_tdtomato@range)[1]-400,end=end,cex.axis = 0.5,cex.title=0.8,background.title="cadetblue2",col.axis="black",col.title="black",col.border.title="black",showSampleNames = TRUE,lwd.title=0.5)

```
### Putting wta and tso tracks together for egfp

```{r}

ylims_unmod <- extendrange(range(c(values(coverage_egfp$unmod_wta), values(coverage_egfp$unmod_tso))))
displayPars(coverage_egfp$unmod_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$unmod_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_unmod_roi <- extendrange(range(c(values(coverage_egfp$unmod_roi_wta), values(coverage_egfp$unmod_roi_tso))))
displayPars(coverage_egfp$unmod_roi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$unmod_roi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rock <- extendrange(range(c(values(coverage_egfp$rock_wta), values(coverage_egfp$rock_tso))))
displayPars(coverage_egfp$rock_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$rock_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rockroi <- extendrange(range(c(values(coverage_egfp$rockroi_wta), values(coverage_egfp$rockroi_tso))))
displayPars(coverage_egfp$rockroi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_egfp$rockroi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

unmod_egfp<-OverlayTrack(trackList=list(coverage_egfp$unmod_wta,coverage_egfp$unmod_tso),legend=TRUE)
unmod_roi_egfp<-OverlayTrack(trackList=list(coverage_egfp$unmod_roi_wta,coverage_egfp$unmod_roi_tso),legend=TRUE)
rock_egfp<-OverlayTrack(trackList=list(coverage_egfp$rock_wta,coverage_egfp$rock_tso),legend=TRUE,ylims=ylims_rock)
rockroi_egfp<-OverlayTrack(trackList=list(coverage_egfp$rockroi_wta,coverage_egfp$rockroi_tso),legend=TRUE)
names(unmod_egfp)<-"unmod"
names(unmod_roi_egfp)<-"unmod_roi"
names(rock_egfp)<-"rock"
names(rockroi_egfp)<-"rockroi"

plotTracks(c(unmod_egfp,unmod_roi_egfp,rock_egfp,rockroi_egfp,gm_egfp),type="histogram",showSampleNames = TRUE,from=start(gm_egfp@range)[1]-400,end=end,cex.axis = 0.5,cex.title=0.8,background.title="cadetblue2",col.axis="black",col.title="black",col.border.title="black",showSampleNames = TRUE,lwd.title=0.5,ylim=c(0,max(ylims_rockroi,ylims_rock,ylims_unmod_roi,ylims_unmod)))

```
### Putting wta and tso tracks together for tdtomato

```{r}

ylims_unmod <- extendrange(range(c(values(coverage_tdtomato$unmod_wta), values(coverage_tdtomato$unmod_tso))))
displayPars(coverage_tdtomato$unmod_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$unmod_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_unmod_roi <- extendrange(range(c(values(coverage_tdtomato$unmod_roi_wta), values(coverage_tdtomato$unmod_roi_tso))))
displayPars(coverage_tdtomato$unmod_roi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$unmod_roi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rock <- extendrange(range(c(values(coverage_tdtomato$rock_wta), values(coverage_tdtomato$rock_tso))))
displayPars(coverage_tdtomato$rock_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$rock_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

ylims_rockroi <- extendrange(range(c(values(coverage_tdtomato$rockroi_wta), values(coverage_tdtomato$rockroi_tso))))
displayPars(coverage_tdtomato$rockroi_wta)<-list(col.histogram="pink",fill.histogram=TRUE)
displayPars(coverage_tdtomato$rockroi_tso)<-list(col.histogram="lightblue",fill.histogram=TRUE)

unmod_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$unmod_wta,coverage_tdtomato$unmod_tso),legend=TRUE)
unmod_roi_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$unmod_roi_wta,coverage_tdtomato$unmod_roi_tso),legend=TRUE)
rock_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$rock_wta,coverage_tdtomato$rock_tso),legend=TRUE)
rockroi_tdtomato<-OverlayTrack(trackList=list(coverage_tdtomato$rockroi_wta,coverage_tdtomato$rockroi_tso),legend=TRUE)

names(unmod_tdtomato)<-"unmod"
names(unmod_roi_tdtomato)<-"unmod_roi"
names(rock_tdtomato)<-"rock"
names(rockroi_tdtomato)<-"rockroi"

plotTracks(c(unmod_tdtomato,unmod_roi_tdtomato,rock_tdtomato,rockroi_tdtomato,gm_tdtomato),type="histogram",showSampleNames = TRUE,from=start(gm_egfp@range)[1]-400,end=end,cex.axis = 0.5,cex.title=0.8,background.title="cadetblue2",col.axis="black",col.title="black",col.border.title="black",showSampleNames = TRUE,lwd.title=0.5,ylim=c(0,max(ylims_rockroi,ylims_rock,ylims_unmod_roi,ylims_unmod)))

```

### Reading .bam files

```{r}

getwd()

bam<-dir(here("00_mixing_experiment"), "*.bam", 
            recursive = TRUE, full.names = TRUE)

bam

names(bam) <- gsub(".bam", "", basename(bam))
names(bam) <- gsub("_subset", "",names(bam))

datadir <- dirname(bam)[1] # folder where the .bw files are 

```


```{r}



```


### Print sessionInfo()

```{r}

sessionInfo()

```



