---
title: "00_mixing_j_mapping_stats"
author: "Mark Robinson, Izaskun Mallona & Giulia Moro"
format: 
  html:
    toc: true
    embed-resources: true
    keep_md: true
---


```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(SingleCellExperiment)
library(scran)
library(ggplot2)
library(dplyr)
library(scater)
library(reshape2)
library(UpSetR)
library(limma)
library(cowplot)
library(pheatmap)
library(googlesheets4)
library(readr)
library(plotROC)
library(here)
library(RColorBrewer)
library(ggnewscale)
library(scales)
library(rmarkdown)
library(khroma)
library(irlba)
library(viridis)
library(gridExtra)
library(quarto)
library(Biostrings)
library(BSgenome)
library(rtracklayer)
library(Matrix)
library(Matrix.utils)
library(edgeR)
library(DESeq2)
library(ggrepel)

packageVersion("Matrix")

```


### Setting up knitr 

```{r}

knitr::opts_chunk$set(fig.width = 2.7,
               fig.height = 2.7,
               fig.align = "center",
               include = TRUE,
               dev = c("png","svg"),
               cache=TRUE,
               cache.lazy = FALSE,
               warning = TRUE,
               message = TRUE)

```

### Themes

```{r}

rock_roi_theme<-function(){
  theme(axis.text = element_text(size=7), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=7),
        aspect.ratio=1, 
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
}

rock_roi_theme_multiple_panels<-function(){
    theme(axis.text = element_text(size=7), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.key.size = unit(0.5, 'cm'),
        legend.title=element_text(size=5,face="bold"),
        legend.text=element_text(size=5),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.border = element_rect(colour = "black", fill = NA))
}

```

### Defining colors

```{r}

myscale<-c(color("muted")(9), gray="darkgrey", 
             black = "black", darkorange = "darkorange",
             yellowgreen = "yellowgreen")

# defining colors for samples and for species

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","unmod_n1","high")

#plot(1:13, col=myscale, pch=19, cex=4)

```


### The format of the rnaqc and bamqc output .txt file is horrible, so transfering data manually (not ideal)

```{r}

qc_stats<-data.frame(matrix(nrow=9,ncol=7))
rownames(qc_stats)<-c("n_mapped_reads",
                      "n_alignments", 
                      "n_aligned_genes",
                      "% exonic",
                      "% intronic",
                      "% intergenic",
                      "mapping_quality (%)",
                      "n_mito_human_bases",
                      "n_mito_mouse_bases")

colnames(qc_stats)<-c("unmod_wta","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso")

qc_stats$unmod_wta<-c(79464700,127015266,45496609,81.86,12.13,6.01,19.76,330626453,175135445)
qc_stats$unmod_roi_wta<-c(61028403,97460697,34441519,80.12,13.96,5.91,19.28,285568718,101339931)
qc_stats$unmod_roi_tso<-c(30623595,41289259,19546682,84.79,12.47,2.74,19.96,493260518,227603853)
qc_stats$rock_wta<-c(64502908,104950805,36268922,81.54,12.04,6.42,19.68,234123637,127966736)
qc_stats$rock_tso<-c(44610313,63685100,29127875,85.24,12.69,2.07,19.18,299489080,133170831)
qc_stats$rockroi_wta<-c(59838042,95051279,34533361,81.87,13.05,5.08,20.25,264061292,100791208)
qc_stats$rockroi_tso<-c(38727236,54463523,25294377,84.78,13.29,1.93,19.43,318747278,99451799)

qc_stats$name<-rownames(qc_stats)

plottable<-melt(qc_stats,id="name")

# for (i in unique(plottable$name)){
#   p<-ggplot(plottable[plottable$name==i,],aes(x=variable,y=value,fill=variable))+
#   geom_bar(stat = "identity")+
#   ggtitle(i)+
#   theme_bw()+
#   rock_roi_theme()+
#   theme(legend.position = "none")+
#   scale_fill_manual(labels=c("unmod_wta","unmod_tso","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso"),
#                      values=alpha(c(myscale[["unmod"]],myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rock"]],myscale[["rockroi"]],myscale[["rockroi"]]),c(1,0.5,1,0.5,1,0.5,1,0.5)))
#   print(p)
# }

```

### Loading data

```{r}

getwd() # after opening the .Rproj file, getwd() should be "/Users/gimoro/figure_making_paper/rock_roi_paper"

rdss <- dir(here("00_mixing_experiment/data/starsolo_stats"), "*.txt", 
            recursive = TRUE, full.names = TRUE)

datadir <- dirname(rdss)[1] # folder where the .rds files are 


samples<-c("unmod_wta","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso")
summary_stats<-list()

for (i in samples){
    dir<-paste0(here("00_mixing_experiment/data/starsolo_stats"),"/",i,"/","Summary.csv")
    summary_stats[[i]]<-read.csv(dir,header=FALSE)
    colnames(summary_stats[[i]])<-c("stats","values")
}

```

### Changing the number of reads based on values calculated in .bam file (see 06_Sankey)

```{r}

summary_stats$unmod_wta$values[1]<-105154707
summary_stats$unmod_roi_wta$values[1]<-94405059
summary_stats$unmod_roi_tso$values[1]<-62804285
summary_stats$rock_wta$values[1]<-100426253
summary_stats$rock_tso$values[1]<-83334432
summary_stats$rockroi_wta$values[1]<-91170527
summary_stats$rockroi_tso$values[1]<-71646657
  
```

### Stats plot

```{r}

for (i in names(summary_stats)){
  summary_stats[[i]]<-summary_stats[[i]][1:9,]
}

# combining everything together

df_summary<-data.frame(matrix(nrow=9,ncol=7))
rownames(df_summary)<-summary_stats[[i]]$stats
colnames(df_summary)<-names(summary_stats)

for(i in names(summary_stats)){
  df_summary[,i]<-summary_stats[[i]]$values
}

df_summary$name<-rownames(df_summary)
plottable<-melt(df_summary,id="name")

# for (i in unique(plottable$name)){
#   p<-ggplot(plottable[plottable$name==i,],aes(x=variable,y=value,fill=variable))+
#   geom_bar(stat = "identity")+
#   ggtitle(i)+
#   theme_bw()+
#   rock_roi_theme()+
#   theme(legend.position = "none")+
#   scale_fill_manual(labels=c("unmod_wta","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso"),
#                      values=alpha(c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rock"]],myscale[["rockroi"]],myscale[["rockroi"]]),c(1,1,0.5,1,0.5,1,0.5)))
# 
#   print(p)
# }

```

### Loading data for Features.stats

```{r}

getwd() # after opening the .Rproj file, getwd() should be "/Users/gimoro/figure_making_paper/rock_roi_paper"

rdss <- dir(here("00_mixing_experiment/data/starsolo_stats"), "*.txt", 
            recursive = TRUE, full.names = TRUE)

datadir <- dirname(rdss)[1] # folder where the .rds files are 

samples<-c("unmod_wta","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso")

feature_stats<-list()

for (i in samples){
    dir<-paste0(here("00_mixing_experiment/data/starsolo_stats"),"/",i,"/","Features.stats")
    feature_stats[[i]]<-read.table(dir,header=FALSE)
    colnames(feature_stats[[i]])<-c("stats","values")
}

```

### Plots for Features.stats

```{r}

# combining everything together

df_features<-data.frame(matrix(nrow=11,ncol=7))
rownames(df_features)<-feature_stats[[i]]$stats
colnames(df_features)<-names(feature_stats)

for(i in names(feature_stats)){
  df_features[,i]<-feature_stats[[i]]$values
}

df_features$name<-rownames(df_features)
plottable<-melt(df_features,id="name")

# for (i in unique(plottable$name)){
#   p<-ggplot(plottable[plottable$name==i,],aes(x=variable,y=value,fill=variable))+
#   geom_bar(stat = "identity")+
#   ggtitle(i)+
#   theme_bw()+
#   rock_roi_theme()+
#   theme(legend.position = "none")+
#   scale_fill_manual(labels=c("unmod_wta","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso"),
#                      values=alpha(c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rock"]],myscale[["rockroi"]],myscale[["rockroi"]]),c(1,1,0.5,1,0.5,1,0.5)))
#   print(p)
# }

```


### Selecting relevant features and making pretty plots

```{r}

merged <- do.call("rbind", list(df_features,df_summary,qc_stats))

selected_merged <- merged[c("n_mapped_reads",
                            "n_alignments",
                            "n_aligned_genes",
                            "n_mito_human_bases",
                            "n_mito_mouse_bases",
                            "Number of Reads",
                            "Reads With Valid Barcodes",
                            "Reads Mapped to Genome: Unique+Multiple",
                            "Reads Mapped to Gene: Unique Gene",
                            "Reads Mapped to Genome: Unique",
                            "noUnmapped",
                            "noNoFeature",
                            "mapping_quality (%)"),]

selected_merged$unmod_wta<-as.numeric(selected_merged$unmod_wta)
selected_merged$unmod_roi_wta<-as.numeric(selected_merged$unmod_roi_wta)
selected_merged$unmod_roi_tso<-as.numeric(selected_merged$unmod_roi_tso)
selected_merged$rock_wta<-as.numeric(selected_merged$rock_wta)
selected_merged$rock_tso<-as.numeric(selected_merged$rock_tso)
selected_merged$rockroi_wta<-as.numeric(selected_merged$rockroi_wta)
selected_merged$rockroi_tso<-as.numeric(selected_merged$rockroi_tso)

for (i in samples){
  selected_merged["normalised_unmapped",i]<-selected_merged["noUnmapped",i]/selected_merged["Number of Reads",i]
  selected_merged["normalised_noNoFeature",i]<-selected_merged["noNoFeature",i]/selected_merged["Number of Reads",i]
}

selected_merged$name<-rownames(selected_merged)

```

### Plotting barplots

```{r}

plottable<-melt(selected_merged,id="name")

for (i in unique(plottable$name)){
  p<-ggplot(plottable[plottable$name==i,],aes(x=variable,y=value,fill=variable))+
  geom_bar(stat = "identity")+
  ggtitle(i)+
  theme_bw()+
  rock_roi_theme()+
  theme(legend.position = "none")+
 scale_fill_manual(labels=c("unmod_wta","unmod_roi_wta","unmod_roi_tso","rock_wta","rock_tso","rockroi_wta","rockroi_tso"),
                     values=alpha(c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rock"]],myscale[["rockroi"]],myscale[["rockroi"]]),c(1,1,0.5,1,0.5,1,0.5)))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  print(p)
}
    
```


### Plotting stacking barplots

```{r}

exo_intro <- merged[c("% intronic","% intergenic","% exonic"),]
exo_intro$unmod_wta<-as.numeric(exo_intro$unmod_wta)
exo_intro$unmod_roi_wta<-as.numeric(exo_intro$unmod_roi_wta)
exo_intro$unmod_roi_tso<-as.numeric(exo_intro$unmod_roi_tso)
exo_intro$rock_wta<-as.numeric(exo_intro$rock_wta)
exo_intro$rock_tso<-as.numeric(exo_intro$rock_tso)
exo_intro$rockroi_wta<-as.numeric(exo_intro$rockroi_wta)
exo_intro$rockroi_tso<-as.numeric(exo_intro$rockroi_tso)

plottable<-melt(exo_intro,id="name")

ggplot(plottable,aes(x=variable,y=value,fill=name))+
  geom_bar(stat = "identity",position="stack")+
  theme_bw()+
  rock_roi_theme()+
  scale_fill_manual(labels=c("% exonic","% intergenic","% intronic"),
                    values=alpha(c(myscale[["high"]],myscale[["high"]],myscale[["high"]]),c(1,0.5,0.2)))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### Session info

```{r}

sessionInfo()

```
