---
title: "00_mixing_experiment_f-paper_figures"
author: "Mark Robinson, Izaskun Mallona & Giulia Moro"
format: 
  html:
    toc: true
    embed-resources: true
    keep_md: true
---

### Load packages, load data 

```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(SingleCellExperiment)
library(scran)
library(ggplot2)
library(dplyr)
library(scater)
library(reshape2)
library(UpSetR)
library(limma)
library(cowplot)
library(pheatmap)
library(readr)
library(plotROC)
library(ggpubr)
library(here)
library(RColorBrewer)
library(ggnewscale)
library(scales)
library(rmarkdown)
library(irlba)
library(viridis)
library(gridExtra)
library(googlesheets4)
library(tools)
library(gridExtra)
library(khroma)

packageVersion("Matrix")

rdss <- dir(here("00_mixing_experiment"), "^mixing.*rds$", 
            recursive = TRUE, full.names = TRUE)

names(rdss) <- gsub("_sce.rds", "", basename(rdss))
names(rdss) <- gsub("mixing_", "", names(rdss))

datadir <- dirname(rdss)[1] # folder where the .rds files are 

### objects generated in report 00_mixing_experiment_c-figures

# unfiltered object 

sces <- mapply(function(u,v)  {
  rds <- readRDS(u)
  rds$sample_id <- v
  rds$sample_id.barcode <- paste0(v,".",colnames(rds))
  # colnames(rowData(rds)) <- c("name", "type", "value","captured")
  g <- grepl("^ENS", rownames(rds))
  rownames(rds)[g] <- paste0(rownames(rds)[g], 
                             "__", rowData(rds)$name[g])
  rowData(rds)$type <- "capture"
  k <- grepl("^ENSG", rownames(rds))
  rowData(rds)$type[k] <- "human_gene"
  k <- grepl("^ENSMUS", rownames(rds))
  rowData(rds)$type[k] <- "mouse_gene"
  k <- grepl("^mt", rowData(rds)$name)
  rowData(rds)$type[k] <- "mouse_mito"
  k <- grepl("^MT", rowData(rds)$name)
  rowData(rds)$type[k] <- "human_mito"
  rds
}, rdss, names(rdss))

sce <- Reduce(cbind, sces)

# filtered object

sce_filt<-readRDS(file.path(datadir, "sce_filt.rds"))

# filtered after doublet removal 

sce_nd<-readRDS(file.path(datadir, "sce_nd.rds"))

```

### Setting up knitr 

```{r}

knitr::opts_chunk$set(fig.width = 2.7,
               fig.height = 2.7,
               fig.align = "center",
               include = TRUE,
               dev = c("png","svg"),
               cache=TRUE,
               cache.lazy = FALSE,
               warning = TRUE,
               message = TRUE)

```

### Themes

```{r}

rock_roi_theme<-function(){
  theme(axis.text = element_text(size=5), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        aspect.ratio=1,
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
}

rock_roi_theme_multiple_panels<-function(){
    theme(axis.text = element_text(size=5), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.key.size = unit(0.5, 'cm'),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.border = element_rect(colour = "black", fill = NA))
}

```

### Defining colors to use

```{r}

myscale<-c(color("muted")(9), gray="darkgrey", 
             black = "black", darkorange = "darkorange",
             yellowgreen = "yellowgreen")

# defining colors for samples and for species

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","low","high")

#plot(1:13, col=myscale, pch=19, cex=4)

```

### Violin plots, colored by species

```{r}

rd <- rowData(sce)
sce <- addPerCellQCMetrics(sce, assay.type = "wta",
                           subsets=list(capture=rd$type=="capture",
                                        human_gene=rd$type=="human_gene",
                                        human_mito=rd$type=="human_mito",
                                        mouse_gene=rd$type=="mouse_gene",
                                        mouse_mito=rd$type=="mouse_mito"))

# adding species

sce$species<-"unknown"
sce$species[which(sce$subsets_mouse_gene_percent>50)]<-"mouse"
sce$species[which(sce$subsets_human_gene_percent>50)]<-"human"

cd <- colData(sce) %>% as.data.frame %>% 
  mutate(total_mito_percent = subsets_mouse_mito_percent+subsets_human_mito_percent)

table(sce$species) # not removing the unknown species yet

ggplot(cd, aes(x=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")), y=detected)) +
        geom_violin()+  
  theme_bw()+
  xlab("")+
  ylab("number of genes")+ 
  rock_roi_theme()+
  geom_jitter(aes(color=sce$species),shape=16, position=position_jitter(0.2),alpha=0.2,size=0.2)+
  geom_boxplot(aes(fill=species),width=0.3,outlier.size = 0.5)+
  labs(colour="species")+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))+
  scale_color_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))+
  scale_fill_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))

ggplot(cd, aes(x=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")), y=total)) +
        geom_violin()+  
  theme_bw()+
  xlab("")+
  ylab("number of UMIs")+ 
  geom_jitter(aes(color=sce$species),shape=16, position=position_jitter(0.2),alpha=0.2,size=0.2)+
  geom_boxplot(aes(fill=species),width=0.3,outlier.size = 0.5)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_color_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))+
  scale_fill_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))

ggplot(cd, aes(x=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")), y=subsets_mouse_mito_percent)) +
        geom_violin()+  
  theme_bw()+
  xlab("")+
  ylab("% mitochondrial transcripts")+ 
  geom_jitter(aes(color=sce$species),shape=16, position=position_jitter(0.2),alpha=0.2,size=0.2)+
  rock_roi_theme()+
  labs(colour="species")+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))+
  scale_color_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))

```

### Plots with UMI vs mito colored by human and mouse 

```{r}

ggplot(cd, aes(x = total, y = total_mito_percent)) +
  geom_point(aes(color=sce$species),size=0.1,alpha=0.5) + 
  scale_x_log10() + 
  scale_y_sqrt() +
  facet_wrap(~factor(sample_id,levels=c("unmod","unmod_roi",
                                        "rock","rockroi"))) + 
  theme_bw()+
  xlab("total number of transcripts")+
  ylab("percent mitochondrial reads")+
  theme(
    axis.text = element_text(size = 15),
    axis.title=element_text(size=20,face="bold"),
    strip.text=element_text(size=15,face="bold"))+
    rock_roi_theme_multiple_panels()+
  geom_hline(yintercept=c(3,25), colour="orange") +
  geom_vline(xintercept=c(2750), colour="orange") +
  labs(colour="species")+
  scale_color_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))+
  scale_fill_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))

```

### Importing annotation file

```{r}

# read annotation info from GS
anno_url <- "https://docs.google.com/spreadsheets/d/1cvldF_VFA7FYuChoR9a4SIpYcZ1aS80yThNT3iP4T00/edit#gid=978414281"
anno_df <- read_sheet(anno_url, "transcript-regions-mark") 
anno_df <- anno_df[anno_df$Experiment=="Mixing",]
regs <- split(anno_df$Name_in_gtf_file, anno_df$region_name)

# compute aggregates for egfp/tdtomato for TSO
ae <- altExp(sce_filt,"tso_ontarget_multi")
cd <- colData(ae)

for(i in 1:length(regs))
  cd[[ names(regs)[i] ]] <- colSums(assay(ae,1)[regs[[i]],,drop=FALSE])

colData(altExp(sce_filt,"tso_ontarget_multi")) <- cd
# ---------------

# compute aggregates for egfp/tdtomato for WTA
ae <- altExp(sce_filt,"wta_ontarget_multi")
cd <- colData(ae)

for(i in 1:length(regs))
  cd[[ names(regs)[i] ]] <- colSums(assay(ae,1)[regs[[i]],,drop=FALSE])

colData(altExp(sce_filt,"wta_ontarget_multi")) <- cd
# ---------------

```

### Score regions tso_multi filt dataset 

```{r}

ae <- altExp(sce_filt,"tso_ontarget_multi",withColData=TRUE)

ae$tso_egfp_detection_score <- 0
ae$tso_egfp_detection_score[ae$egfp_tx>0] <- 1
ae$tso_egfp_detection_score[ae$egfp_cds>0] <- 2
ae$tso_egfp_detection_score[ae$egfp_roi>0] <- 3
ae$tso_egfp_detection_score[ae$egfp_roi>0] <- 4

ae$tso_tdtom_detection_score<-0
ae$tso_tdtom_detection_score[ae$tdtom_tx>0] <- 1
ae$tso_tdtom_detection_score[ae$tdtom_cds>0] <- 2
ae$tso_tdtom_detection_score[ae$tdtom_roi1>0] <- 3
ae$tso_tdtom_detection_score[ae$tdtom_roi2>0] <- 3
ae$tso_tdtom_detection_score[ae$tdtom_roi2>0 & ae$tdtom_roi1>0] <- 4

table(ae$tso_egfp_detection_score)
table(ae$tso_tdtom_detection_score)

altExp(sce_filt,"tso_ontarget_multi") <- ae

```

### Score regions wta_multi filt dataset 

```{r}

ae <- altExp(sce_filt,"wta_ontarget_multi",withColData=TRUE)

ae$wta_egfp_detection_score <- 0
ae$wta_egfp_detection_score[ae$egfp_tx>0] <- 1
ae$wta_egfp_detection_score[ae$egfp_cds>0] <- 2
ae$wta_egfp_detection_score[ae$egfp_roi>0] <- 3
ae$wta_egfp_detection_score[ae$egfp_roi>0] <- 4

ae$wta_tdtom_detection_score <- 0
ae$wta_tdtom_detection_score[ae$tdtom_tx>0] <- 1
ae$wta_tdtom_detection_score[ae$tdtom_cds>0] <- 2
ae$wta_tdtom_detection_score[ae$tdtom_roi1>0] <- 3
ae$wta_tdtom_detection_score[ae$tdtom_roi2>0] <- 3
ae$wta_tdtom_detection_score[ae$tdtom_roi2>0 & ae$tdtom_roi1>0] <- 4

table(ae$wta_egfp_detection_score)
table(ae$wta_tdtom_detection_score)

altExp(sce_filt,"wta_ontarget_multi") <- ae

```

### Splitting mouse vs human filt dataset and removing cells with unknown species

```{r}

sce_filt$species <- "unknown"
sce_filt$species[sce_filt$subsets_mouse_gene_percent>50] <- "mouse"
sce_filt$species[sce_filt$subsets_human_gene_percent>50] <- "human"

cd <- colData(sce_filt) %>% as.data.frame %>% 
  mutate(total_mito_percent = subsets_mouse_mito_percent+subsets_human_mito_percent)

table(sce_filt$species,
      altExp(sce_filt,"tso_ontarget_multi",withColData=TRUE)$tso_egfp_detection_score)
table(sce_filt$species,
      altExp(sce_filt,"tso_ontarget_multi",withColData=TRUE)$tso_tdtom_detection_score)

```

### Barnyard plot mouse vs human (unknow removed and filtered dataset)

```{r}

sce_filt$sample_id<-gsub("mixing_", "", sce_filt$sample_id)

cd<-colData(sce_filt) %>% as.data.frame

ggplot(cd, aes(x = subsets_mouse_gene_sum,y = subsets_human_gene_sum,color=sce_filt$species))+
  geom_point(alpha = 0.2,size=0.5) +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~factor(sample_id, levels = c("unmod", "unmod_roi", "rock", "rockroi"))) +
  theme_bw() +
  xlab("mouse gene sum") +
  ylab("gene sum")+
  ggtitle(" ")+
  rock_roi_theme_multiple_panels()+
  guides(colour = guide_legend(override.aes = list(alpha = 1)))+
  scale_color_manual(values=c("human"=myscale[["human"]],"mouse"=myscale[["mouse"]],"unknown"=myscale[["unknown"]]))+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))+
  labs(colour="species")
  
```

### Removing cells with unknown species

```{r}

sce_filt <- sce_filt[, !sce_filt$species== "unknown"]

```


### Barnyard plot with CDS detection

```{r}

combined<-colData(sce_filt) %>% as.data.frame

combined$wta_tdtom_detection_score<-altExp(sce_filt,"wta_ontarget_multi")$wta_tdtom_detection_score
combined$wta_egfp_detection_score<-altExp(sce_filt,"wta_ontarget_multi")$wta_egfp_detection_score
combined$tso_tdtom_detection_score<-altExp(sce_filt,"tso_ontarget_multi")$tso_tdtom_detection_score
combined$tso_egfp_detection_score<-altExp(sce_filt,"tso_ontarget_multi")$tso_egfp_detection_score

combined$color<-myscale[["doublets"]] # these will be doublets

combined$color[which(combined$tso_egfp_detection_score==0 & combined$tso_tdtom_detection_score==0)]<-myscale[["no_detection"]] # no detection of either
combined$color[which(combined$tso_egfp_detection_score==1 & combined$tso_tdtom_detection_score>0)]<-myscale[["no_detection"]]# no detection of either
combined$color[which(combined$tso_egfp_detection_score>0 & combined$tso_tdtom_detection_score==1)]<-myscale[["no_detection"]] # no detection of either

combined$color[which(combined$tso_egfp_detection_score==2 & combined$tso_tdtom_detection_score<=1)]<-myscale[["egfp_cds"]] # rois are part of the cds
combined$color[which(combined$tso_egfp_detection_score==3 & combined$tso_tdtom_detection_score<=1)]<-myscale[["egfp_cds"]] # rois are part of the cds
combined$color[which(combined$tso_egfp_detection_score==4 & combined$tso_tdtom_detection_score<=1)]<-myscale[["egfp_cds"]] # rois are part of the cds

combined$color[which(combined$tso_egfp_detection_score<=1 & combined$tso_tdtom_detection_score==2)]<-myscale[["tdtom_cds"]] # rois are part of the cds
combined$color[which(combined$tso_egfp_detection_score<=1 & combined$tso_tdtom_detection_score==3)]<-myscale[["tdtom_cds"]] # rois are part of the cds
combined$color[which(combined$tso_egfp_detection_score<=1 & combined$tso_tdtom_detection_score==4)]<-myscale[["tdtom_cds"]]# rois are part of the cds

ggplot(combined, aes(
  x = subsets_mouse_gene_sum,
  y = subsets_human_gene_sum,
  color = color 
)) +
  geom_point(alpha = 0.4,size=0.5) +
  scale_x_log10() + scale_y_log10() +
  facet_wrap(~factor(sample_id, levels = c("unmod", "unmod_roi", "rock", "rockroi"))) +
  theme_bw() +
  xlab("mouse gene sum") +
  ylab("human gene sum") +
  rock_roi_theme_multiple_panels()+
  scale_color_identity(
    name = "Legend Title",
    guide = "legend")+
  labs(colour="CDS_detection")+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))+
  scale_color_manual(labels=c("egfp_cds","tdtom_cds","doublets","no_detection"), 
                     values=c("#117733","#AA4499","black","darkgray"))

```

### TSO vs WTA plot, colored by on target tso_ontarget_multi

```{r,fig.width=10,fig.height=2.7}

sce_nd$sample_id<-gsub("mixing_", "", sce_nd$sample_id)

rds <- rowData(sce_nd)

tso_qc <- perCellQCMetrics(sce_nd, assay.type = "tso_off_and_ontarget_unique",
                           subsets=list(capture=rds$type=="capture",
                                        human_gene=rds$type=="human_gene",
                                        human_mito=rds$type=="human_mito",
                                        mouse_gene=rds$type=="mouse_gene",
                                        mouse_mito=rds$type=="mouse_mito"))

df <- data.frame(tso_total = tso_qc$total,
                 tso_capture_sum = tso_qc$subsets_capture_sum,
                 tso_mito_percent = tso_qc$subsets_mouse_mito_percent+
                   tso_qc$subsets_human_mito_percent,
                 tso_detected = tso_qc$detected,
                 wta_total = sce_nd$total,
                 wta_capture_sum = sce_nd$subsets_capture_sum,
                 wta_mito_percent = sce_nd$subsets_mouse_mito_percent+
                   sce_nd$subsets_human_mito_percent,
                 sample_id = sce_nd$sample_id)

# specifying regions since don't want to include the regulatory regions of egfp and tdtomato

ae_tso <- altExp(sce_nd,"tso_ontarget_multi")
ae_wta <- altExp(sce_nd,"wta_ontarget_multi")

df$tso_capture_sum<-colSums(assay(ae_tso,1)[regs$egfp_tx,,drop=FALSE])+colSums(assay(ae_tso,1)[regs$tdtom_tx,,drop=FALSE])
df$wta_capture_sum<-colSums(assay(ae_wta,1)[regs$egfp_tx,,drop=FALSE])+colSums(assay(ae_wta,1)[regs$tdtom_tx,,drop=FALSE])

df$log_1p_tso_ontarget_multi <- log1p(df$tso_capture_sum)

p <- list()

for (i in  c("unmod", "unmod_roi", "rock", "rockroi")){
     p[[i]]<-ggplot(df[df$sample_id==i,], aes(x = tso_total, y = wta_total, color = log_1p_tso_ontarget_multi)) +
          geom_point(alpha=0.2,size=0.5) +
          scale_x_log10(limit=c(1,30000)) + 
          scale_y_log10(limit=c(2000,40000)) +
          theme_bw() +
          xlab("total TSO UMIs") +
          ylab("total WTA UMIs") +
          labs(title=i)+
          rock_roi_theme_multiple_panels()+
          theme(legend.key.width=unit(0.3,"cm"))+
          labs(colour="TSO ontarget")+
          scale_colour_gradient2(limits=c(min(df$log_1p_tso_ontarget_multi),max(df$log_1p_tso_ontarget_multi)),
                         low=myscale[["no_detection"]],
                         mid=myscale[["no_detection"]],
                         high=myscale[[i]],
                         midpoint=mean(df$log_1p_tso_ontarget_multi))
}

grid.arrange(grobs=p,nrow=1)

```

### Plot with just the total TSO UMI

```{r}

ggplot(df, aes(x=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")), y=tso_total,fill=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")))) +
  geom_violin(show.legend = FALSE)+  
  theme_bw()+
  xlab("")+
  ylab("total TSO UMI")+ 
  scale_y_log10()+
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_fill_manual(labels=c("unmod","unmod_roi","rock","rockroi"),
                     values=c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rockroi"]]))

```



### Same for mito

```{r}

ggplot(df, aes(x=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")), y=tso_mito_percent,fill=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")))) +
  geom_violin(show.legend = FALSE)+  
  theme_bw()+
  xlab("")+
  ylab("% mitochondrial transcripts")+ 
  scale_y_log10()+
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_fill_manual(labels=c("unmod","unmod_roi","rock","rockroi"),
                     values=c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rockroi"]]))

```

### Same for genes

```{r}

ggplot(df, aes(x=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")), y=tso_detected,fill=factor(sample_id,levels=c("unmod","unmod_roi","rock","rockroi")))) +
  geom_violin(show.legend = FALSE)+  
  theme_bw()+
  xlab("")+
  ylab("total TSO genes")+ 
  scale_y_log10()+
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_fill_manual(labels=c("unmod","unmod_roi","rock","rockroi"),
                     values=c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rockroi"]]))

```

### False positive vs true positive

```{r,fig.height=3,fig.width=6}

exprs_nd <- assay(altExp(sce_nd,"tso_ontarget_multi",withColData=TRUE),1)

sort(rowSums(exprs_nd, na.rm = TRUE), decreasing = TRUE) -> sorted

df <- data.frame(mouse = (sce_nd$species == "mouse")+0,
                 human = (sce_nd$species == "human")+0,
                 sample = sce_nd$sample_id,
                 as.matrix(t(exprs_nd[names(sorted),])))

dfm <- melt(df, id.vars = c("human","mouse","sample"))

human_targets <- grep("tdtomato", names(sorted), value = TRUE)

p <- ggplot(dfm %>% dplyr::filter(variable %in% c("roi_1_tdtomato", "roi_2_tdtomato","rock_tdtomato")),
                      aes(m = value, d = human, colour=sample, group=sample)) + 
  geom_roc(size = 0.3, alpha = .8,
           position = position_dodge2(width=.001),
           labels=FALSE,
           linealpha=0.5,
           size.point=0.3) +
  facet_grid(~variable) +
  xlim(0, .018) +
  ggtitle("Human targets") +
  theme_bw()+
  rock_roi_theme_multiple_panels()+
  scale_color_manual(labels=c("unmod","unmod_roi","rock","rockroi"), 
                     values=c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rockroi"]]))+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))

mouse_targets <- grep("egfp", names(sorted), value = TRUE)

q <- ggplot(dfm %>% filter(variable %in% c("rock_egfp","roi_egfp")),
                      aes(m = value, d = mouse, colour=sample, group=sample)) + 
  geom_roc(size = 0.3, alpha = .8,
           position = position_dodge2(width=.001),
           linealpha=0.5,
           labels=FALSE,
           size.point=0.3) +
  facet_grid(~variable) +
  xlim(0, .018) +
  ggtitle("Mouse targets") +
  theme_bw()+
  rock_roi_theme_multiple_panels()+
  theme(plot.title=element_text(margin=margin(t=0.1,unit="cm")))+
  scale_color_manual(labels=c("unmod","unmod_roi","rock","rockroi"), 
                     values=c(myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rockroi"]]))+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))

grid.arrange(p,q)

```

### Library size per cell WTA vs TSO (previous plots from Izaskun)

```{r,fig.width=10,fig.height=2.7}

rds <- rowData(sce_nd)

tso_qc <- perCellQCMetrics(sce_nd, assay.type = "tso_off_and_ontarget_unique",
                           subsets=list(capture=rds$type=="capture",
                                        human_gene=rds$type=="human_gene",
                                        human_mito=rds$type=="human_mito",
                                        mouse_gene=rds$type=="mouse_gene",
                                        mouse_mito=rds$type=="mouse_mito"))

df <- data.frame(tso_total = tso_qc$total,
                 tso_capture_sum = tso_qc$subsets_capture_sum,
                 tso_mito_percent = tso_qc$subsets_mouse_mito_percent+
                   tso_qc$subsets_human_mito_percent,
                 wta_total = sce_nd$total,
                 wta_capture_sum = sce_nd$subsets_capture_sum,
                 wta_mito_percent = sce_nd$subsets_mouse_mito_percent+
                   sce_nd$subsets_human_mito_percent,
                 sample_id = sce_nd$sample_id)

ae_tso <- altExp(sce_nd,"tso_ontarget_multi")
ae_wta <- altExp(sce_nd,"wta_ontarget_multi")

df$tso_capture_sum<-colSums(assay(ae_tso,1)[regs$egfp_tx,,drop=FALSE])+colSums(assay(ae_tso,1)[regs$tdtom_tx,,drop=FALSE])
df$wta_capture_sum<-colSums(assay(ae_wta,1)[regs$egfp_tx,,drop=FALSE])+colSums(assay(ae_wta,1)[regs$tdtom_tx,,drop=FALSE])

samples<-c("unmod","unmod_roi","rock","rockroi")

p <- list()

for (i in samples){
     p[[i]]<-ggplot(df[df$sample_id==i,], aes(x = wta_capture_sum,
                    y = tso_capture_sum,
                    color = wta_total)) +
      geom_point(alpha = 0.5,size=0.5) +
      geom_smooth(colour="black") +
      scale_x_continuous(trans='log1p',limits=c(0,max(df$wta_capture_sum))) +
      scale_y_continuous(trans='log1p',limits=c(0,max(df$tso_capture_sum))) +
      theme_bw() +
      labs(title=i) +
      xlab('on-target WTA # UMIs') +
      ylab('on-target TSO # UMIs')+
      rock_roi_theme_multiple_panels()+
      theme(legend.position = "left",
            legend.key.width=unit(0.3,"cm"))+
      scale_colour_gradient2(limits=c(min(df$wta_total),max(df$wta_total)),
                         low=myscale[["no_detection"]],
                         mid=myscale[[i]],
                         high=myscale[[i]],
                         midpoint=max(df$wta_total))
     
      p[[i]]<-ggExtra::ggMarginal(p[[i]],type = "density")
}

grid.arrange(grobs=p,nrow=1)

```


### Barplot with % positive cells, ORF vs tx in same plot for mouse and human

```{r}

df$egfp_capture_sum_cds<-colSums(assay(ae_tso,1)[regs$egfp_cds,,drop=FALSE])+colSums(assay(ae_wta,1)[regs$egfp_cds,,drop=FALSE])
df$tdtom_capture_sum_cds<-colSums(assay(ae_tso,1)[regs$tdtom_cds,,drop=FALSE])+colSums(assay(ae_wta,1)[regs$tdtom_cds,,drop=FALSE])

df$egfp_capture_sum<-colSums(assay(ae_tso,1)[regs$egfp_tx,,drop=FALSE])+colSums(assay(ae_wta,1)[regs$egfp_tx,,drop=FALSE])
df$tdtom_capture_sum<-colSums(assay(ae_tso,1)[regs$tdtom_tx,,drop=FALSE])+colSums(assay(ae_wta,1)[regs$tdtom_tx,,drop=FALSE])

df$species<-sce_nd$species

mouse<-df[df$species=="mouse",]
human<-df[df$species=="human",]

positive_cells<-data.frame(matrix(nrow=8,ncol=2))

colnames(positive_cells)<-c("mouse_egfp","human_tdtomato")

rows<-c()
for (i in unique(df$sample_id)){
  rows<-c(rows,paste0(i,"_CDS"),paste0(i,"_tx"))
}

rownames(positive_cells)<-rows

for (i in unique(mouse$sample_id)){
  positive_cells[paste0(i,"_CDS"),1]<-length(which(mouse[mouse$sample_id==i,]$egfp_capture_sum_cds>0))/
    length(mouse[mouse$sample_id==i,]$egfp_capture_sum_cds)*100
  positive_cells[paste0(i,"_tx"),1]<-length(which(mouse[mouse$sample_id==i,]$egfp_capture_sum>0))/
    length(mouse[mouse$sample_id==i,]$egfp_capture_sum)*100
  positive_cells[paste0(i,"_CDS"),2]<-length(which(human[human$sample_id==i,]$tdtom_capture_sum_cds>0))/
    length(human[human$sample_id==i,]$tdtom_capture_sum_cds)*100
    positive_cells[paste0(i,"_tx"),2]<-length(which(human[human$sample_id==i,]$tdtom_capture_sum>0))/
      length(human[human$sample_id==i,]$tdtom_capture_sum)*100
}

plottable<-melt(positive_cells)

plottable$sample<-rep(rownames(positive_cells),2)

plottable$hjs<-rep(c(2,2,2,2,-0.5,2,-0.5,2),2)

samples<-c("unmod_CDS","unmod_tx","unmod_roi_CDS","unmod_roi_tx","rock_CDS","rock_tx","rockroi_CDS","rockroi_tx")

ggplot(data=plottable,aes(y=value,x=variable,fill=factor(sample,levels=samples)))+
  geom_bar(stat="identity",width = 0.8, position = position_dodge(width = 0.9),alpha=c(rep(c(1,0.5),8)))+
  xlab(" ")+
  ylab("positive cells (%)")+
  theme_bw()+
  ylim(0,100)+
  rock_roi_theme()+
  geom_text(aes(label=format(round(value,2),nsmall=2),hjust=hjs,vjust=0.5), 
            position=position_dodge(width=0.9), 
            size=2,
            angle = 90)+
  labs(fill="samples",fill = NULL)+
  scale_fill_manual(labels=samples, 
                     values=alpha(c(myscale[["unmod"]],myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rock"]],myscale[["rockroi"]],myscale[["rockroi"]]),c(rep(c(1,0.5),4))))+
  guides(colour = guide_legend(override.aes = list(size=3)))

```

### Barplot with opposite cell type / fluorescent transcript

```{r}

colnames(positive_cells)<-c("mouse_tdtomato","human_egfp")

rows<-c()
for (i in unique(df$sample_id)){
  rows<-c(rows,paste0(i,"_CDS"),paste0(i,"_tx"))
}

rownames(positive_cells)<-rows

for (i in unique(mouse$sample_id)){
  positive_cells[paste0(i,"_CDS"),1]<-length(which(mouse[mouse$sample_id==i,]$tdtom_capture_sum_cds>0))/
    length(mouse[mouse$sample_id==i,]$tdtom_capture_sum_cds)*100
  positive_cells[paste0(i,"_tx"),1]<-length(which(mouse[mouse$sample_id==i,]$tdtom_capture_sum>0))/
    length(mouse[mouse$sample_id==i,]$tdtom_capture_sum)*100
  positive_cells[paste0(i,"_CDS"),2]<-length(which(human[human$sample_id==i,]$egfp_capture_sum_cds>0))/
    length(human[human$sample_id==i,]$egfp_capture_sum_cds)*100
    positive_cells[paste0(i,"_tx"),2]<-length(which(human[human$sample_id==i,]$egfp_capture_sum>0))/
      length(human[human$sample_id==i,]$egfp_capture_sum)*100
}

plottable<-melt(positive_cells)

plottable$sample<-rep(rownames(positive_cells),2)

plottable$hjs<-rep(c(-0.5,2,-0.5,2,-0.5,2,-0.5,2),2)

samples<-c("unmod_CDS","unmod_tx","unmod_roi_CDS","unmod_roi_tx","rock_CDS","rock_tx","rockroi_CDS","rockroi_tx")

ggplot(data=plottable,aes(y=value,x=variable,fill=factor(sample,levels=samples)))+
  geom_bar(stat="identity",width = 0.8, position = position_dodge(width = 0.9),alpha=c(rep(c(1,0.5),8)))+
  xlab(" ")+
  ylab("positive cells (%)")+
  theme_bw()+
  ylim(0,100)+
  rock_roi_theme()+
  geom_text(aes(label=format(round(value,2),nsmall=2),hjust=hjs,vjust=0.5), 
            position=position_dodge(width=0.9), 
            size=2,
            angle = 90)+
  labs(fill="samples",fill = NULL)+
  scale_fill_manual(labels=samples, 
                     values=alpha(c(myscale[["unmod"]],myscale[["unmod"]],myscale[["unmod_roi"]],myscale[["unmod_roi"]],myscale[["rock"]],myscale[["rock"]],myscale[["rockroi"]],myscale[["rockroi"]]),c(rep(c(1,0.5),4))))+
  guides(colour = guide_legend(override.aes = list(size=3)))

```

### Session info

```{r}

sessionInfo()

```

