---
title: "03_first_mixing_experiment_e-tso_data"
author: "Mark Robinson, Izaskun Mallona & Giulia Moro"
format: 
  html:
    toc: true
    embed-resources: true
    keep_md: true
---

### Load packages

```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(SingleCellExperiment)
library(scran)
library(ggplot2)
library(dplyr)
library(scater)
library(reshape2)
library(UpSetR)
library(limma)
library(cowplot)
library(pheatmap)
library(googlesheets4)
library(readr)
library(plotROC)
library(here)
library(RColorBrewer)
library(ggnewscale)
library(scales)
library(rmarkdown)
library(khroma)
library(irlba)
library(viridis)
library(gridExtra)
library(quarto)
library(Biostrings)
library(BSgenome)
library(rtracklayer)
library(GenomicRanges)
library(GenomicFeatures)
library(memes)
library(universalmotif)
library(plyranges)

```

### Setting up knitr 

```{r}

knitr::opts_chunk$set(fig.width = 2.7,
               fig.height = 2.7,
               fig.align = "center",
               include = TRUE,
               dev = c("png","svg"),
               cache=TRUE,
               cache.lazy = FALSE,
               warning = TRUE,
               message = TRUE)

```

### Themes

```{r}

rock_roi_theme<-function(){
  theme(axis.text = element_text(size=7), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=7),
        aspect.ratio=1, 
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
}

rock_roi_theme_multiple_panels<-function(){
    theme(axis.text = element_text(size=7), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.key.size = unit(0.5, 'cm'),
        legend.title=element_text(size=5,face="bold"),
        legend.text=element_text(size=5),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.border = element_rect(colour = "black", fill = NA))
}

```

### Defining colors

```{r}

myscale<-c(color("muted")(9), gray="darkgrey", 
             black = "black", darkorange = "darkorange",
             yellowgreen = "yellowgreen")

# defining colors for samples and for species

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","unmod_n1","high")

#plot(1:13, col=myscale, pch=19, cex=4)

```

### Load data

```{r}

getwd() # after opening the .Rproj file, getwd() should be "/Users/gimoro/figure_making_paper/rock_roi_paper"

rdss <- dir(here("03_first_mixing_experiment"), "^mixing.*rds", 
            recursive = TRUE, full.names = TRUE)

names(rdss) <- gsub("_sce.rds", "", basename(rdss))
names(rdss) <- gsub("mixing_", "", names(rdss))

datadir <- dirname(rdss)[1] # folder where the .rds files are 

sce_nd <- readRDS(file.path(datadir, "sce_nd.rds")) # running analysis on filtered object

```

### Comparison top 100 WTA of unmod_n1 and TSO unmod_n1

```{r,fig.height=7,fig.width=5.4}

sce_nd$sample_id <- gsub("mixing_", "", sce_nd$sample_id)

n1 <- sce_nd[,sce_nd$sample_id==c("unmod_n1")]
n1_wta <- assay(n1,"wta")
n1_tso <- assay(n1,"tso_off_and_ontarget_unique")

top_100_wta<-sort(rowSums(n1_wta),decreasing=TRUE)[1:100]
top_100_tso<-sort(rowSums(n1_tso),decreasing=TRUE)[1:100]

length(intersect(names(top_100_wta),names(top_100_tso))) # 50

wta_for_top_100_tso<-rowSums(n1_wta[names(top_100_tso),])
top_100<-data.frame(wta=wta_for_top_100_tso,tso=sort(top_100_tso,decreasing=TRUE))

wta_for_top_100_tso<-rowSums(n1_wta[names(top_100_tso),])

top_100<-data.frame(wta=wta_for_top_100_tso,tso=top_100_tso)
top_100<-top_100[order(top_100$tso,decreasing=TRUE),]

tso_only<-names(top_100_tso)
tso_only[!tso_only %in% names(top_100_wta)] <- "" # this way in heatmap remove gene name

# order tso_only based on the order of the expressed genes

# for all top 100, only labelling the TSO only 

pheatmap(top_100,
         show_rownames = TRUE,
         show_colnames = TRUE, 
         scale="none",
         fontsize_row = 5,
         fontsize_col = 7,
         fontsize = 5,
         treeheight_row = 0,
         treeheight_col = 0,
         cellwidth=15,
         labels_row = tso_only,
         cluster_rows=F)

```

### From this checking the % mt of TSO vs WTA

```{r}

rds <- rowData(sce_nd)

tso_qc <- perCellQCMetrics(n1, assay.type = "tso_off_and_ontarget_unique",
                           subsets=list(capture=rds$type=="capture",
                                        human_gene=rds$type=="human_gene",
                                        human_mito=rds$type=="human_mito",
                                        mouse_gene=rds$type=="mouse_gene",
                                        mouse_mito=rds$type=="mouse_mito"))

df_tso <- data.frame(tso_total = tso_qc$total,
                 tso_detected = tso_qc$detected,
                 tso_capture_sum = tso_qc$subsets_capture_sum,
                 tso_mito_percent = tso_qc$subsets_mouse_mito_percent+
                   tso_qc$subsets_human_mito_percent)

wta_qc <- perCellQCMetrics(n1, assay.type = "wta",
                           subsets=list(capture=rds$type=="capture",
                                        human_gene=rds$type=="human_gene",
                                        human_mito=rds$type=="human_mito",
                                        mouse_gene=rds$type=="mouse_gene",
                                        mouse_mito=rds$type=="mouse_mito"))

df_wta <- data.frame(wta_total = wta_qc$total,
                 wta_detected = wta_qc$detected,
                 wta_capture_sum = wta_qc$subsets_capture_sum,
                 wta_mito_percent = wta_qc$subsets_mouse_mito_percent+
                   wta_qc$subsets_human_mito_percent)

df<-data.frame(df_wta$wta_mito_percent,df_tso$tso_mito_percent)
colnames(df)<-c("wta","tso")
plottable<-melt(df)

ggplot(plottable,aes(x=variable,y=value,fill=variable)) +
  geom_violin(show.legend = FALSE) +  
  theme_bw() +
  xlab("") +
  ylab("percent mitochondrial transcripts")+ 
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE) +
  rock_roi_theme() +
  scale_fill_manual(labels=c("wta","tso"),
                     values=alpha(c(myscale[["unmod_n1"]],myscale[["unmod_n1"]]),c(1,0.3)))

```

## Detecting TSO motif in the TSO specific highly expressed transcripts

### First of all need to get the transcript sequences for the genes of interest --> need the .gtf and .fa file 

```{r}

gtf <- rtracklayer::import(file.path(datadir,"mixing.gtf"))

library("BSgenome.Hsapiens.UCSC.hg38") # object: Hsapiens, different patch than the one we used
library("BSgenome.Mmusculus.UCSC.mm10") # object: Mmusculus, same patch as the one we used

seqnames(Hsapiens)<-paste0("human_",seqnames(Hsapiens))
seqnames(Mmusculus)<-paste0("mouse_",seqnames(Mmusculus))

```

### Matching to mouse genome TSO, - strand is the TSO. Searching for 6 / 5 / 4. Just do mito 

```{r}

# rules based on manually viewing some examples

#TSO sequence "TATGCGTAGTAGGTATG"
#rev TSO: "CATACCTACTACGCATA"

dnaseq1 <- DNAString("GGTATG")
dnaseq2 <- DNAString("GTATG")
dnaseq3 <- DNAString("TATG")

match_mouse1<-vmatchPattern(dnaseq1, Mmusculus, max.mismatch=1, min.mismatch=0, exclude = seqnames(Mmusculus)[-22])
match_mouse2<-vmatchPattern(dnaseq2, Mmusculus, max.mismatch=1, min.mismatch=0, exclude = seqnames(Mmusculus)[-22])
match_mouse3<-vmatchPattern(dnaseq3, Mmusculus, max.mismatch=1, min.mismatch=0, exclude = seqnames(Mmusculus)[-22])

```

### Matching to mito rev TSO for + strand 

```{r}

# rev comp TSO: CATACCTACTAC

revdnaseq1 <- DNAString("CATACC")
revdnaseq2 <- DNAString("CATAC")
revdnaseq3 <- DNAString("CATA")

revmatch_mouse1<-vmatchPattern(revdnaseq1, Mmusculus, max.mismatch=1, min.mismatch=0, exclude = seqnames(Mmusculus)[-22])
revmatch_mouse2<-vmatchPattern(revdnaseq2, Mmusculus, max.mismatch=1, min.mismatch=0, exclude = seqnames(Mmusculus)[-22])
revmatch_mouse3<-vmatchPattern(revdnaseq3, Mmusculus, max.mismatch=1, min.mismatch=0, exclude = seqnames(Mmusculus)[-22])

```

### Matching to human genome TSO 

```{r}

match_human1<-vmatchPattern(dnaseq1, Hsapiens, max.mismatch=1, min.mismatch=0, exclude = seqnames(Hsapiens)[-25])
match_human2<-vmatchPattern(dnaseq2, Hsapiens, max.mismatch=1, min.mismatch=0, exclude = seqnames(Hsapiens)[-25])
match_human3<-vmatchPattern(dnaseq3, Hsapiens, max.mismatch=1, min.mismatch=0, exclude = seqnames(Hsapiens)[-25])

```

### Matching to human genome rev TSO

```{r}

revmatch_human1<-vmatchPattern(revdnaseq1, Hsapiens, max.mismatch=1, min.mismatch=0, exclude = seqnames(Hsapiens)[-25])
revmatch_human2<-vmatchPattern(revdnaseq2, Hsapiens, max.mismatch=1, min.mismatch=0, exclude = seqnames(Hsapiens)[-25])
revmatch_human3<-vmatchPattern(revdnaseq3, Hsapiens, max.mismatch=1, min.mismatch=0, exclude = seqnames(Hsapiens)[-25])

```

### Subsetting to + genes with rev --> 

```{r}

gtf_plus<-gtf[which(strand(gtf)=="+"),]

revmouse1<-subsetByOverlaps(revmatch_mouse1,gtf_plus)
revmouse2<-subsetByOverlaps(revmatch_mouse2,gtf_plus)
revmouse3<-subsetByOverlaps(revmatch_mouse3,gtf_plus)

revhuman1<-subsetByOverlaps(revmatch_human1,gtf_plus)
revhuman2<-subsetByOverlaps(revmatch_human2,gtf_plus)
revhuman3<-subsetByOverlaps(revmatch_human3,gtf_plus)

```

### Subsetting to - genes with TSO

```{r}

gtf_minus<-gtf[which(strand(gtf)=="-"),]

mouse1<-subsetByOverlaps(match_mouse1,gtf_minus)
mouse2<-subsetByOverlaps(match_mouse2,gtf_minus)
mouse3<-subsetByOverlaps(match_mouse3,gtf_minus)

human1<-subsetByOverlaps(match_human1,gtf_minus)
human2<-subsetByOverlaps(match_human2,gtf_minus)
human3<-subsetByOverlaps(match_human3,gtf_minus)

match_6<-list(mouse1,human1,revmouse1,revmouse1)
match_6<-do.call(c, as(match_6, "GRangesList"))

match_5<-list(mouse2,human2,revmouse2,revhuman2)
match_5<-do.call(c, as(match_5, "GRangesList"))

match_4<-list(mouse3,human3,revmouse3,revhuman3)
match_4<-do.call(c, as(match_4, "GRangesList"))

```

### GRanges to bed

```{r}

bed_6<-data.frame(seqnames=seqnames(match_6),
  starts=start(match_6)-1,
  ends=end(match_6),
      strands=strand(match_6))

bed_5<-data.frame(seqnames=seqnames(match_5),
  starts=start(match_5)-1,
  ends=end(match_5),
      strands=strand(match_5))

bed_4<-data.frame(seqnames=seqnames(match_4),
  starts=start(match_4)-1,
  ends=end(match_4),
      strands=strand(match_4))

write.table(bed_6, file=file.path(datadir,"match_6.bed"), quote=F, sep="\t", row.names=F, col.names=F)
write.table(bed_5, file=file.path(datadir,"match_5.bed"), quote=F, sep="\t", row.names=F, col.names=F)
write.table(bed_4, file=file.path(datadir,"match_4.bed"), quote=F, sep="\t", row.names=F, col.names=F)

```

### Runnning motif finder

```{r}

# run for genes of interest
# no rev needed 
# removing the last 5 bp of the tso at the 5' end 

tso <-"TATGCGTAGTAGGTATG"
revtso <-"CATACCTACTACGCATA"

tso_motif.matrix<-matrix(c(0.4/3,0.4/3,0.6,  0.4/3,0.4/3,0.6,  0,0,0,1,0,0, #A
                           0.4/3,0.4/3,0.4/3,0.4/3,0.4/3,0.4/3,0,0,0,0,0,0, #C
                           0.6,  0.4/3,0.4/3,0.6,  0.4/3,0.4/3,1,1,0,0,0,1,    #G
                           0.4/3,0.6,  0.4/3,0.4/3,0.6,  0.4/3,0,0,1,0,1,0),#T
                           nrow=4,byrow=TRUE)

tso_motif<-create_motif(tso_motif.matrix, alphabet="DNA",name = "TSO",pseudocount=1,nsites=17,type="PPM")

tso_motif<-motif_rc(tso_motif)
view_motifs(tso_motif,use.type="PPM")

```

### Searching for motif in mouse in mitochondrial genes

```{r}

set<-list()
for (i in seqnames(Mmusculus)[1:22]){
  set[[i]]<-getSeq(Mmusculus,i)
}

set<-DNAStringSet(set)[c(22)] # for just mt 

mouse_plus<-runFimo(set,tso_motif,skip_matched_sequence=TRUE,thresh=0.0005) # need to change thresh otherwise doesn't show any results

```

### Searching for motif in human in mitochondrial genes

```{r}

set<-list()
for (i in seqnames(Hsapiens)[1:25]){
  set[[i]]<-getSeq(Hsapiens,i)
}

set<-DNAStringSet(set)[c(25)] # for just mt 

human_plus<-runFimo(set,tso_motif,skip_matched_sequence=TRUE,thresh=0.0005)

```

### Need to get information about the strand in the .gtf but still maintain the match information

```{r}

human_genes<-find_overlaps_directed(human_plus,gtf)
mouse_genes<-find_overlaps_directed(mouse_plus,gtf)

human_genes<-human_genes[human_genes$type=="gene",] # otherwise get multiple lines for each match (exon / transcript / gene)
mouse_genes<-mouse_genes[mouse_genes$type=="gene",]

```

### Granges to bed

```{r}

bed_mouse_genes<-data.frame(seqnames=seqnames(mouse_genes),
  starts=start(mouse_genes)-1,
  ends=end(mouse_genes),
      strands=strand(mouse_genes))

bed_human_genes<-data.frame(seqnames=seqnames(human_genes),
  starts=start(human_genes)-1,
  ends=end(human_genes),
      strands=strand(human_genes))
  
write.table(bed_mouse_genes, file=file.path(datadir,"mouse_mt.bed"), quote=F, sep="\t", row.names=F, col.names=F)
write.table(bed_human_genes, file=file.path(datadir,"human_MT.bed"), quote=F, sep="\t", row.names=F, col.names=F)

```

### Running for whole genome (for 6)

```{r}

set<-list()
for (i in seqnames(Mmusculus)[1:22]){
  set[[i]]<-getSeq(Mmusculus,i)
}
set<-DNAStringSet(set)
mouse_plus<-runFimo(set,tso_motif,skip_matched_sequence=TRUE,thresh=0.0005) 

set<-list()
for (i in seqnames(Hsapiens)[1:25]){
  set[[i]]<-getSeq(Hsapiens,i)
}
set<-DNAStringSet(set)
human_plus<-runFimo(set,tso_motif,skip_matched_sequence=TRUE,thresh=0.0005)

human_genes<-find_overlaps_directed(gtf,human_plus)
mouse_genes<-find_overlaps_directed(gtf,mouse_plus)

human_genes$combined_name<-paste0(human_genes$gene_id,"__",human_genes$gene_name)
mouse_genes$combined_name<-paste0(mouse_genes$gene_id,"__",mouse_genes$gene_name)

human_genes<-human_genes[human_genes$type=="gene",]
mouse_genes<-mouse_genes[mouse_genes$type=="gene",]

df_mouse<-data.frame(presence_match=as.vector(table(mouse_genes$combined_name)))
rownames(df_mouse)<-names(table(mouse_genes$combined_name))

df_human<-data.frame(presence_match=as.vector(table(human_genes$combined_name)))
rownames(df_human)<-names(table(human_genes$combined_name))

```

### Checking in general if TSO pattern match == higher gene expression

```{r,fig.height=2.7,fig.width=5.4}

# need to get the gene expression for the transcripts with the TSO motif

n1_tso_sums<-data.frame(tso_count=rowSums(n1_tso))
n1_tso_sums$wta_count<-as.data.frame(rowSums(n1_wta))[rownames(n1_tso_sums),]

# adding gene length to dataframe

gene_gtf<-gtf[gtf$type=="gene",]
gene_gtf$width<-width(ranges(gene_gtf))
gene_gtf_df<-as.data.frame(gene_gtf)
gene_gtf_df$combined<-paste0(gene_gtf_df$gene_id,"__",gene_gtf_df$gene_name)
gene_gtf_width<-as.data.frame(gene_gtf_df$width)
rownames(gene_gtf_width)<-gene_gtf_df$combined

# combining

n1_tso_sums$gene_length<-gene_gtf_width[rownames(n1_tso_sums),]

# adding presence match

n1_tso_sums$presence_match<-0

for (i in intersect(rownames(df_mouse),rownames(n1_tso_sums))){
   n1_tso_sums[i,"presence_match"]<-1
}
 
for (i in intersect(rownames(df_human),rownames(n1_tso_sums))){
   n1_tso_sums[i,"presence_match"]<-1
}

table(n1_tso_sums$presence_match)

# exporting object

write.csv(n1_tso_sums, paste0(here("03_first_mixing_experiment"),"/data/","n1_tso_sums.csv"), row.names=TRUE)

ggplot(data=n1_tso_sums,aes(x=as.character(presence_match),y=tso_count,fill=as.character(presence_match)))+
  geom_boxplot(outlier.size=1)+
  geom_jitter(shape=16, position=position_jitter(0.1),size=0.1,alpha=0.2)+
  scale_y_log10()+
  theme_bw()+
  xlab("presence_match")+
  rock_roi_theme()+
  theme(legend.position = "none")+
  scale_fill_manual(labels=c("0","1"),
    values=c(myscale[["rockroi"]],"#8856A7"))

```

### Checking if genes detected in TSO overlap with genes detected in WTA

```{r}

n1_wta<-n1_wta[names(which(rowSums(n1_wta)>10)),]
n1_tso<-n1_tso[names(which(rowSums(n1_tso)>10)),]

### compare expression level of WTA only compared to both

wta_only<-data.frame(wta_count=rowSums(n1_wta[setdiff(rownames(n1_wta),rownames(n1_tso)),]))
wta_only$name<-"wta_only"
both_wta_tso<-data.frame(wta_count=rowSums(n1_wta[intersect(rownames(n1_wta),rownames(n1_tso)),]))
both_wta_tso$name<-"both"

plottable<-melt(rbind(wta_only,both_wta_tso))

ggplot(data=plottable,aes(x=name,y=value,fill=name))+
  geom_boxplot(outlier.size=1)+
  geom_jitter(shape=16, position=position_jitter(0.1),size=0.1,alpha=0.2)+
  theme_bw()+
  scale_y_log10()+
  xlab("")+
  rock_roi_theme()+
  ylab("wta_count")+
  theme(legend.position = "none")+
  scale_fill_manual(labels=c("wta_only","both"),
                    values=alpha(c(myscale[["unknown"]],myscale[["unknown"]]),c(1,0.3)))

```

### Same with the ones only detected in TSO

```{r}

n1_wta<-n1_wta[names(which(rowSums(n1_wta)>10)),]
n1_tso<-n1_tso[names(which(rowSums(n1_tso)>10)),]

### compare expression level of WTA only compared to both

tso_only<-data.frame(wta_count=rowSums(n1_tso[setdiff(rownames(n1_tso),rownames(n1_wta)),]))
tso_only$name<-"tso_only"
both_wta_tso<-data.frame(wta_count=rowSums(n1_tso[intersect(rownames(n1_wta),rownames(n1_tso)),]))
both_wta_tso$name<-"both"

plottable<-melt(rbind(tso_only,both_wta_tso))

ggplot(data=plottable,aes(x=name,y=value,fill=name))+
  geom_boxplot(outlier.size=1)+
  geom_jitter(shape=16, position=position_jitter(0.1),size=0.1,alpha=0.2)+
  theme_bw()+
  scale_y_log10()+
  xlab("")+
  rock_roi_theme()+
  ylab("tso_count")+
  theme(legend.position = "none")+
  scale_fill_manual(labels=c("tso_only","both"),
                    values=alpha(c(myscale[["human"]],myscale[["human"]]),c(1,0.3)))

```

### Plot from Izaskun: (log expression) vs (log expression rank) for each gene in tso, and color the dot by having a TSO pattern or not

```{r}

## log expression for wta
## log rank

ranks<-n1_tso_sums[,c("tso_count","presence_match")]
ranks$log10_exp<-log10(ranks$tso_count+1) # added 1 to avoid inf

ranks<-ranks[order(ranks$log10_exp,decreasing=TRUE),]
ranks$rank<-seq(1,length(rownames(ranks)),by=1)

ggplot(ranks,aes(x=rank,y=log10_exp,color=as.factor(presence_match)))+
  geom_point(alpha=0.3)+
  theme_bw()+
  rock_roi_theme()+
  labs(color="presence match")+
  scale_color_manual(labels=c("0","1"),
    values=c(myscale[["rockroi"]],"#8856A7"))+
  guides(colour = guide_legend(override.aes = list(alpha = 1,size=1)))

```

### Checking correlation between tso_count and presence_match

```{r}

model_presence<-lm(tso_count ~ presence_match,data=n1_tso_sums)

# checking if the presence counts

summary(model_presence) # small correlation but very low R-squared

```

### Session info

```{r}

sessionInfo()

```
