---
title: "01_pdgfra_experiment_g-paper_figures"
format:
  html:
    toc: true
    embed-resources: true
editor_options: 
  chunk_output_type: console
---


```{r,warning=FALSE, message=FALSE}

rm(list=ls())

library(SingleCellExperiment)
library(scran)
library(ggplot2)
library(dplyr)
library(scater)
library(reshape2)
library(UpSetR)
library(limma)
library(Rsamtools)
library(cowplot)
library(pheatmap)
library(readr)
library(plotROC)
library(ggpubr)
library(here)
library(RColorBrewer)
library(ggnewscale)
library(scales)
library(rmarkdown)
library(irlba)
library(viridis)
library(gridExtra)
library(googlesheets4)
library(tools)
library(gridExtra)
library(khroma)
library(dittoSeq)
library(rtracklayer)
library(Gviz)
library(cowplot)
library(GenomicFeatures)
library(GenomicRanges)
library(corrplot)
library(ggrastr)

datadir <- file.path(here("01_pdgfra_experiment"), "data")

rdss <- dir(here("01_pdgfra_experiment"), "*_sce.rds", 
            recursive = TRUE, full.names = TRUE)

names(rdss) <- gsub("_sce.rds", "", basename(rdss))
names(rdss)<-c("m_pdgfra_rockroi","u_pdgfra_rockroi","m_pdgfra_unmod")

datadir <- dirname(rdss)[1]

sces <- mapply(function(u,v)  {
  rds <- readRDS(u)
  rds$sample_id <- v
  rds$sample_id.barcode <- paste0(v,".",colnames(rds))
  colnames(rowData(rds)) <- c("name", "type", "value")
  g <- grepl("^ENS", rownames(rds))
  rownames(rds)[g] <- paste0(rownames(rds)[g], 
                             "__", rowData(rds)$name[g])
  rowData(rds)$gene_type <- "capture"
  k <- grepl("^ENSG", rownames(rds))
  rowData(rds)$gene_type[k] <- "human_gene"
  k <- grepl("^ENSMUS", rownames(rds))
  rowData(rds)$gene_type[k] <- "mouse_gene"
  k <- grepl("^mt", rowData(rds)$name)
  rowData(rds)$gene_type[k] <- "mouse_mito"
  k <- grepl("^MT", rowData(rds)$name)
  rowData(rds)$gene_type[k] <- "human_mito"
  rds
}, rdss, names(rdss))

sce <- Reduce(cbind, sces)

```

### Setting up knitr 

```{r}

knitr::opts_chunk$set(fig.width = 2.7,
               fig.height = 2.7,
               fig.align = "center",
               include = TRUE,
               dev = c("png","svg"),
               cache=TRUE,
               cache.lazy = FALSE,
               warning = TRUE,
               message = TRUE)

```

### Themes

```{r}

rock_roi_theme<-function(){
  theme(axis.text = element_text(size=5), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        aspect.ratio=1,
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
}

rock_roi_theme_multiple_panels<-function(){
    theme(axis.text = element_text(size=5), 
        plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=7,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        legend.key.size = unit(0.5, 'cm'),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5,
        panel.border = element_rect(colour = "black", fill = NA))
}

```

### Defining colors to use (also adding pdgfra unimodal)

```{r}

myscale<-c(color("muted")(9), gray="darkgrey", 
             black = "black", darkorange = "darkorange",
             yellowgreen = "yellowgreen")

# defining colors for samples and for species

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","low","pdgfra_unimodal")

#plot(1:13, col=myscale, pch=19, cex=4)

```

### QC plots

```{r,fig.height=2.7,fig.width=5.4}

sce<-sce[,sce$sample_id==c("m_pdgfra_rockroi","m_pdgfra_unmod")]
sce$sample_id<-gsub("m_pdgfra_","",sce$sample_id)

table(sce$sample_id)

rd <- rowData(sce)
sce <- addPerCellQCMetrics(sce, assay.type = "wta",
                           subsets=list(capture=rd$gene_type=="capture",
                                        mouse_gene=rd$gene_type=="mouse_gene",
                                        mouse_mito=rd$gene_type=="mouse_mito"))

cd <- colData(sce) %>% as.data.frame

plot1<-ggplot(cd, aes(x = total,
               y = detected,
               colour = subsets_mouse_mito_percent)) +
  geom_point(size=0.8,alpha=0.5,stroke=NA) +
  scale_x_log10() + 
  scale_y_log10() +
  facet_wrap(~factor(sample_id,levels=c("unmod","rockroi")))+
  theme_bw()+
  theme(
    axis.text = element_text(size = 15),
    axis.title=element_text(size=20,face="bold"),
    strip.text=element_text(size=15,face="bold"))+
    rock_roi_theme_multiple_panels()+
  labs(colour="percent mito")

plot1<-ggrastr::rasterize(plot1, layers='Point', dpi=1000)

plot1

ggsave(file=file.path(datadir,"qc1.svg"), plot=plot1, width=3.2, height=2.7)


plot2<-ggplot(cd, aes(x = total, y = subsets_mouse_mito_percent)) +
  geom_point(size=0.8,alpha=0.5,stroke=NA) + 
  scale_x_log10() + scale_y_sqrt() +
  facet_wrap(~factor(sample_id,levels=c("unmod","rockroi")))+
  theme_bw()+
  geom_hline(yintercept=c(1,75), colour="orange") +
  geom_vline(xintercept=c(2000), colour="orange") +
  theme_bw()+
  theme(
    axis.text = element_text(size = 15),
    axis.title=element_text(size=20,face="bold"),
    strip.text=element_text(size=15,face="bold"))+
    rock_roi_theme_multiple_panels()+
  ylab("percent mito")

plot2<-ggrastr::rasterize(plot2, layers='Point', dpi=1000)

plot2

ggsave(file=file.path(datadir,"qc2.svg"), plot=plot2, width=3.2, height=2.7)

  
plot3<-ggplot(cd, aes(x = detected, y = subsets_mouse_mito_percent)) +
  geom_point(size=0.8,alpha=0.5,stroke=NA) + 
  scale_x_log10() + scale_y_sqrt() +
  facet_wrap(~factor(sample_id,levels=c("unmod","rockroi")))+
  geom_hline(yintercept=c(1,75), colour="orange") +
  geom_vline(xintercept=c(800), colour="orange") +
  theme_bw()+
  theme(
    axis.text = element_text(size = 15),
    axis.title=element_text(size=20,face="bold"),
    strip.text=element_text(size=15,face="bold"))+
    rock_roi_theme_multiple_panels()+
  ylab("percent mito")

plot3<-ggrastr::rasterize(plot3, layers='Point', dpi=1000)

plot3

ggsave(file=file.path(datadir,"qc3.svg"), plot=plot3, width=3.2, height=2.7)

rm(sce)
rm(sces)

mean(cd[which(cd$sample_id=="unmod"),]$subsets_mouse_mito_percent)
mean(cd[which(cd$sample_id=="rockroi"),]$subsets_mouse_mito_percent)

mean(cd[which(cd$sample_id=="unmod"),]$detected)
mean(cd[which(cd$sample_id=="rockroi"),]$detected)

```

### QC on TSO data (without doublets for consistency)

```{r,fig.height=2.7,fig.width=2.7}

sces<-readRDS(file.path(datadir,"sce_qc-filtered-nodoublets-clustered-annotated.rds"))

rds <- rowData(sces)
tso_qc <- perCellQCMetrics(sces, assay.type = "tso_off_and_ontarget_unique",
                           subsets=list(capture=rds$gene_type=="capture",
                                        mouse_gene=rds$gene_type=="mouse_gene",
                                        mouse_mito=rds$gene_type=="mouse_mito"))

all(rownames(tso_qc)==colnames(sces))

df <- data.frame(tso_total = tso_qc$total,
                 tso_capture_sum = tso_qc$subsets_capture_sum,
                 tso_mito_percent = tso_qc$subsets_mouse_mito_percent,
                 tso_detected = tso_qc$detected,
                 wta_total = sces$total,
                 wta_capture_sum = sces$subsets_capture_sum,
                 wta_mito_percent = sces$subsets_mouse_mito_percent,
                 sample_id = sces$sample_id)

plot4<-ggplot(df, aes(x=factor(sample_id,levels=c("unmod","rockroi")), y=tso_total,fill=factor(sample_id,levels=c("unmod","rockroi")))) +
  geom_violin(show.legend = FALSE)+  
  theme_bw()+
  xlab("")+
  ylab("total TSO UMI")+ 
  scale_y_log10()+
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_fill_manual(labels=c("unmod","rockroi"),
                     values=c(myscale[["unmod"]],myscale[["rockroi"]]))

plot4<-ggrastr::rasterize(plot4, layers='Point', dpi=1000)

plot4

ggsave(file=file.path(datadir,"tsoumis.svg"), plot=plot4, width=2.7, height=2.7)


plot5<-ggplot(df, aes(x=factor(sample_id,levels=c("unmod","rockroi")), y=tso_mito_percent,fill=factor(sample_id,levels=c("unmod","rockroi")))) +
  geom_violin(show.legend = FALSE)+  
  theme_bw()+
  xlab("")+
  ylab("percent mito transcripts TSO")+ 
  scale_y_log10()+
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_fill_manual(labels=c("unmod","rockroi"),
                     values=c(myscale[["unmod"]],myscale[["rockroi"]]))

plot5<-ggrastr::rasterize(plot5, layers='Point', dpi=1000)

plot5

ggsave(file=file.path(datadir,"tsogenes.svg"), plot=plot5, width=2.7, height=2.7)

plot6<-ggplot(df, aes(x=factor(sample_id,levels=c("unmod","rockroi")), y=tso_detected,fill=factor(sample_id,levels=c("unmod","rockroi")))) +
  geom_violin(show.legend = FALSE)+  
  theme_bw()+
  xlab("")+
  ylab("total TSO genes")+ 
  scale_y_log10()+
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5,size=0.2,show.legend = FALSE)+
  rock_roi_theme()+
  labs(colour="species")+
  scale_fill_manual(labels=c("unmod","rockroi"),
                     values=c(myscale[["unmod"]],myscale[["rockroi"]]))

plot6<-ggrastr::rasterize(plot6, layers='Point', dpi=1000)

plot6

ggsave(file=file.path(datadir,"tsogenes.svg"), plot=plot6, width=2.7, height=2.7)

```

### Dataset for GEO

```{r}

sce<-readRDS(file.path(datadir, "uni_multi_filtered_doublet_removed.rds"))

rr<-as.data.frame(colData(sces[,sces$sample_id=="rockroi"]))
un<-as.data.frame(colData(sces[,sces$sample_id=="unmod"]))

rr$barcode<-gsub("pdgfra_rockroi_multimodal.","",rr$sample_id.barcode)
un$barcode<-gsub("pdgfra_unmod_unimodal.","",un$sample_id.barcode)

rr$broad_annotation<-"Mesenchymal"
rr$broad_annotation[!grepl("Crypt",rr$annotation)]<-"Epithelial"

rownames(rr)<-rr$barcode
rownames(un)<-un$barcode

un$broad_annotation<-"Mesenchymal"
un$broad_annotation[!grepl("Crypt",un$annotation)]<-"Epithelial"

rr_uni<-as.data.frame(colData(sce[,sce$sample_id=="pdgfra_rockroi_unimodal"]))
rr_uni$barcode<-gsub("pdgfra_rockroi_unimodal.","",rr_uni$sample_id.barcode)
rr_uni<-rr_uni[intersect(rr_uni$barcode,rr$barcode),]

rr_small<-rr[intersect(rr$barcode,rr_uni$barcode),]

rr_uni$annotation<-rr_small$annotation
rr_uni$broad_annotation<-rr_small$broad_annotation

rr_uni<-rr_uni[,c("broad_annotation","annotation")]

write.csv(rr_uni,"rockroi_unimodal_annotation.csv")

rr<-rr[,c("broad_annotation","annotation")]
un<-un[,c("broad_annotation","annotation")]

write.csv(rr,"rockroi_multimodal_annotation.csv")
write.csv(un,"unmod_multimodal_annotation.csv")

```

### Clusters for sample id 

```{r,fig.width=5.4,fig.height=2.7}

plot7<-plotUMAP(sces, colour_by = "sample_id",point_alpha=0.5,point_size=0.1) + 
  facet_wrap(~factor(colour_by,levels=c("unmod","rockroi")))+
  rock_roi_theme_multiple_panels()+
  scale_color_manual(values = c(myscale[["unmod"]],myscale[["rockroi"]]))


plot7<-ggrastr::rasterize(plot7, layers='Point', dpi=1000)

plot7

ggsave(file=file.path(datadir,"umapsamples.svg"), plot=plot7, width=5.4, height=2.7)

```

### Clusters for total / percent mT

```{r,fig.width=5.4,fig.height=2.7}

plot8<-plotUMAP(sces, colour_by = "total", point_size = 0.1,point_alpha=0.5) + 
  facet_wrap(~factor(sces$sample_id,levels=c("unmod","rockroi")))+
  rock_roi_theme_multiple_panels()+
  scale_colour_gradient2(low="lightgrey",
                         mid="lightgrey",
                         high=myscale[["low"]])+
  labs(colour="total")

plot8<-ggrastr::rasterize(plot8, layers='Point', dpi=1000)

plot8

ggsave(file=file.path(datadir,"umapgenesmt.svg"), plot=plot8, width=3.2, height=2.7)


plot9<-plotUMAP(sces, colour_by = "subsets_mouse_mito_percent", point_size = 0.1,point_alpha=0.5) + 
  facet_wrap(~factor(sces$sample_id,levels=c("unmod","rockroi")))+
  rock_roi_theme_multiple_panels()+
  scale_colour_gradient2(low="lightgrey",
                         mid="lightgrey",
                         high=myscale[["low"]])+
  labs(colour="percent mt")

plot9<-ggrastr::rasterize(plot9, layers='Point', dpi=1000)

plot9

ggsave(file=file.path(datadir,"umappercentmt.svg"), plot=plot9, width=3.2, height=2.7)

```

### Clusters with cell types

```{r,fig.height=5.4,fig.width=5.4}

names(myscale)<-unique(sces$annotation)

plot10<-plotUMAP(sces, colour_by="annotation",
         point_size = 0.1, point_alpha = 0.5) + 
  facet_wrap(~factor(sces$sample_id,levels=c("unmod","rockroi")))+
  rock_roi_theme_multiple_panels()+
  scale_color_manual(values=myscale)+
  guides(colour = guide_legend(override.aes = list(size=1)))

plot10<-ggrastr::rasterize(plot10, layers='Point', dpi=1000)

plot10

ggsave(file=file.path(datadir,"umapannotation.svg"), plot=plot10, width=5.4, height=5.4)

```

### Dotplot with markers for cell types

```{r,fig.height=5.4,fig.width=8.1}

markers <- read_sheet("https://docs.google.com/spreadsheets/d/1cvldF_VFA7FYuChoR9a4SIpYcZ1aS80yThNT3iP4T00/edit#gid=1756474524",
                      sheet = "markers-pdgfra")

short_markers<-as.vector(sapply(strsplit(markers$Marker,"__"),"[",2))
short_sces<-sces[markers$Marker,]
rownames(short_sces)<-short_markers

assay(short_sces)

plot11<-dittoDotPlot(short_sces,
             short_markers,
             group.by="annotation",
             min.color="grey", 
             max.color="darkorange",
             scale=FALSE,
             size=2.5)+
  rock_roi_theme()

plot11<-ggrastr::rasterize(plot11, layers='Point', dpi=1000)

plot11

ggsave(file=file.path(datadir,"markers.svg"), plot=plot11, width=5.2, height=5.2)

rm(short_sces)

```

### Epithelial vs mesenchymal

```{r,fig.width=2.7,fig.height=5.4}

epi <- rownames(sces) %in%
  markers$Marker[markers$Broad_cell_type=="Epithelial"]

mes <- rownames(sces) %in%
  markers$Marker[markers$Broad_cell_type=="Mesenchymal"]

df <- cbind(df, epithelial_broad = colMeans(logcounts(sces)[epi,]),
            mesenchymal_broad = colMeans(logcounts(sces)[mes,]))

df$annotation<-sces$annotation

cols <- unique(markers$Cell_type)

for(i in 1:length(cols)) {
  ms <- markers$Marker[markers$Cell_type==cols[i]]
  broad <- markers$Broad_cell_type[markers$Cell_type==cols[i]]
  df[[paste0(broad[1],
             "__",cols[i])]] <- colMeans(logcounts(sces)[ms,,drop=FALSE])
}

with(df, table(mesenchymal_broad > .35, epithelial_broad > .35))

plot12<-ggplot(df, aes(x = mesenchymal_broad, y = epithelial_broad, colour = annotation)) +
  geom_jitter(width=.01, height = .01,size=0.8,stroke=NA) +
  geom_hline(yintercept = .35, colour="orange", size = 1) +
  geom_vline(xintercept = .35, colour="orange", size = 1) +
  theme_bw()+
  rock_roi_theme()+
  scale_color_manual(values=myscale)

plot12<-ggrastr::rasterize(plot12, layers='Point', dpi=1000)

plot12

ggsave(file=file.path(datadir,"markers.svg"), plot=plot12, width=2.7, height=2.7)

```

### Correlation plot

```{r,fig.width=5.4,fig.height=5.4}

table(sces$annotation,sces$sample_id)

with(colData(sces) %>% as.data.frame,
     table(annotation, sample_id))

inds <- split(seq_len(ncol(sces)), 
              paste0(sces$sample_id, "__", sces$annotation))

set.seed(1976)
inds <- lapply(inds, function(u) {
  sample(u, ifelse(length(u)>=11, 11, length(u)),
         replace = FALSE)
})

lcounts <- logcounts(sces)

pbs <- sapply(inds, function(u) rowMeans(lcounts[,u]))

m <- rowMeans(pbs)
v <- rowVars(pbs)

smoothScatter(m, v)

keep <- m > .1 & v > .5

length(keep)

pheatmap(cor(pbs[keep,],method="pearson"))

corrplot(cor(pbs[keep,],method="pearson")[c("rockroi__Crypt_bottom_1",
                                            "rockroi__Crypt_bottom_2",
                                            "rockroi__Crypt_top",
                                            "rockroi__Enterocyte_1",
                                            "rockroi__Enterocyte_2",
                                            "rockroi__Enterocyte_3",
                                            "rockroi__Enterocyte_4",
                                            "rockroi__Enterocyte_5",
                                            "rockroi__Enteroendocrine",
                                            "rockroi__Secretory_Goblet",
                                            "rockroi__Secretory_Reg4",
                                            "rockroi__Stem/TA",
                                            "rockroi__Tuft"),
                                          c("unmod__Crypt_bottom_1",
                                            "unmod__Crypt_bottom_2",
                                            "unmod__Crypt_top",
                                            "unmod__Enterocyte_1",
                                            "unmod__Enterocyte_2",
                                            "unmod__Enterocyte_3",
                                            "unmod__Enterocyte_4",
                                            "unmod__Enterocyte_5",
                                            "unmod__Enteroendocrine",
                                            "unmod__Secretory_Goblet",
                                            "unmod__Secretory_Reg4",
                                            "unmod__Stem/TA",
                                            "unmod__Tuft")],order="hclust",
         addCoef.col = 'darkorange',
         number.cex=0.4,
         tl.col="black") 

```

### Corrplot dataset Brügger et al.

```{r,fig.width=3.3,fig.height=3.3}

bruegger<-readRDS(paste0(here("01_pdgfra_experiment"),"/data/DSSuntreated_new.rds"))

inds <- split(seq_len(ncol(bruegger)),bruegger$ident)

lcounts <- logcounts(bruegger)

pbs <- sapply(inds, function(u) rowMeans(lcounts[,u]))

m <- rowMeans(pbs)
v <- rowVars(pbs)

smoothScatter(m, v)

keep <- m > .1 & v > .5

corrplot(cor(pbs[keep,],method="pearson"),order="hclust",
         addCoef.col = 'darkorange',
         number.cex=0.3,
         tl.col="black",
         tl.cex = 0.5) 

```

### Junction analysis

```{r,fig.width=2.3,fig.height=2.3}

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","low","pdgfra_unimodal")

datadir <- file.path(here("01_pdgfra_experiment"), 
                     "data")

# read in SummarizedExperiment with the junction counts

b_tso <- readRDS(file = file.path(datadir,
                              "junctions", "tso_df-annotated-junctions.rds"))

b_wta <- readRDS(file = file.path(datadir,
                              "junctions", "wta_df-annotated-junctions.rds"))

# Same plots for TSO and WTA separately, start with TSO

(aa <- ggplot(b_tso %>% dplyr::filter(grepl("egfp", region)),
                    aes(x = celltype, y = perc_expressed,
                        fill= factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed,
                        colour = factor(sample,levels=c("unmod","rockroi"))))+
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), 
            colour = "black",
            position = position_dodge(width = 0.9),
            size=2,
            vjust=17) +
  facet_wrap(~region, nrow = 1) +
  theme_bw() +
  rock_roi_theme_multiple_panels() +
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none"))+
  xlab("")

# same for WTA

(aa <- ggplot(b_wta[which(b_wta$region_name=="[roi_egfp] egfp"),],
                    aes(x = celltype, y = perc_expressed,
                        fill= factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed,
                        colour = factor(sample,levels=c("unmod","rockroi"))))+
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), 
            colour = "black",
            position = position_dodge(width = 0.9),
            size=2,
            vjust=17) +
  facet_wrap(~region, nrow = 1) +
  theme_bw() +
  rock_roi_theme_multiple_panels() +
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none"))+
  xlab("")

```

### ROIs in cell types

```{r,fig.height=2.3,fig.width=4}

# For TSO

junctions1<-b_tso %>% dplyr::filter(grepl("roi_[456]", region))
regions<-c("[roi_4] chr5:75166771-75167837","[roi_5] chr5:75167967-75170495","[roi_6] chr5:75170666-75170762") # filtering canonical regions
junctions1<-junctions1[which(junctions1$region %in% regions),]

bb <- ggplot(junctions1,
                    aes(x = celltype, y = perc_expressed, fill = factor(sample,levels=c("unmod","rockroi")),
                        colour = factor(sample,levels=c("unmod","rockroi")), group = factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=0.5) +
  facet_wrap(~region, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=3,face="bold"),
        )+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none")

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"barplot1_tso.svg"), plot=bb, width=4, height=2.2)


# For WTA

junctions1<-b_wta %>% dplyr::filter(grepl("roi_[456]", region))
regions<-c("[roi_4] chr5:75166771-75167837","[roi_5] chr5:75167967-75170495","[roi_6] chr5:75170666-75170762") # filtering canonical regions
junctions1<-junctions1[which(junctions1$region %in% regions),]

bb <- ggplot(junctions1,
                    aes(x = celltype, y = perc_expressed, fill = factor(sample,levels=c("unmod","rockroi")),
                        colour = factor(sample,levels=c("unmod","rockroi")), group = factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=0.5) +
  facet_wrap(~region, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=3,face="bold"),
        )+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none")

bb

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

ggsave(file=file.path(datadir,"barplot1_wta.svg"), plot=bb, width=4, height=2.2)
  
```

### Other regions

```{r,fig.height=2.3,fig.width=4}

# For TSO

b_tso %>% dplyr::filter(grepl("roi_[456]", region), 
             celltype=="Epithelial", sample=="rockroi")

junctions2<-b_tso %>% dplyr::filter(grepl("roi_1[3456]", region))
regions2<-c("[roi_13] chr5:75180272-75180999","[roi_14] chr5:75181109-75181522","[roi_15] chr5:75181675-75182976","[roi_16] chr5:75183142-75185514") # filtering canonical regions
junctions2<-junctions2[which(junctions2$region %in% regions2),]

(cc<- ggplot(junctions2,
                    aes(x = celltype, y = perc_expressed, fill = factor(sample,levels=c("unmod","rockroi")),
                        colour = factor(sample,levels=c("unmod","rockroi")), group = factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=0.5) +
  facet_wrap(~region, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=3,face="bold"),
        )+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none"))

cc<-ggrastr::rasterize(cc, layers='Point', dpi=1000)

cc

ggsave(file=file.path(datadir,"barplot2_TSO.svg"), plot=cc, width=4.3, height=2.2)

# For WTA

b_wta %>% dplyr::filter(grepl("roi_[456]", region), 
             celltype=="Epithelial", sample=="rockroi")

junctions2<-b_wta %>% dplyr::filter(grepl("roi_1[3456]", region))
regions2<-c("[roi_13] chr5:75180272-75180999","[roi_14] chr5:75181109-75181522","[roi_15] chr5:75181675-75182976","[roi_16] chr5:75183142-75185514") # filtering canonical regions
junctions2<-junctions2[which(junctions2$region %in% regions2),]

(cc<- ggplot(junctions2,
                    aes(x = celltype, y = perc_expressed, fill = factor(sample,levels=c("unmod","rockroi")),
                        colour = factor(sample,levels=c("unmod","rockroi")), group = factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=0.5) +
  facet_wrap(~region, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=3,face="bold"),
        )+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none"))

cc<-ggrastr::rasterize(cc, layers='Point', dpi=1000)

cc

ggsave(file=file.path(datadir,"barplot2_WTA.svg"), plot=cc, width=4.3, height=2.2)

```

### With region name

```{r,fig.height=2.3,fig.width=4}

# For TSO

b_tso$region_name<-gsub(" chr5","",b_tso$region_name)

junctions1<-b_tso %>% dplyr::filter(grepl("roi_[456]", region))
regions<-c("[roi_4] chr5:75166771-75167837","[roi_5] chr5:75167967-75170495","[roi_6] chr5:75170666-75170762") # filtering canonical regions
junctions1<-junctions1[which(junctions1$region %in% regions),]

names(myscale)<-unique(sces$annotation)

bb <- ggplot(junctions1,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), 
                        y = perc_expressed, 
                        fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab("")

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"barplot3_tso.svg"), plot=bb, width=4, height=1.9)



# For WTA

b_wta$region_name<-gsub(" chr5","",b_wta$region_name)

junctions1<-b_wta %>% dplyr::filter(grepl("roi_[456]", region))
regions<-c("[roi_4] chr5:75166771-75167837","[roi_5] chr5:75167967-75170495","[roi_6] chr5:75170666-75170762") # filtering canonical regions
junctions1<-junctions1[which(junctions1$region %in% regions),]

names(myscale)<-unique(sces$annotation)

bb <- ggplot(junctions1,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), 
                        y = perc_expressed, 
                        fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab("")

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"barplot3_wta.svg"), plot=bb, width=4, height=1.9)

```

### With region name other regions

```{r,fig.height=2.3,fig.width=4}

# For TSO

junctions2<-b_tso %>% dplyr::filter(grepl("roi_1[3456]", region))
regions2<-c("[roi_13] chr5:75180272-75180999","[roi_14] chr5:75181109-75181522","[roi_15] chr5:75181675-75182976","[roi_16] chr5:75183142-75185514") # filtering canonical regions
junctions2<-junctions2[which(junctions2$region %in% regions2),]

b_tso$region_name<-gsub(" chr5","",b_tso$region_name)

names(myscale)<-unique(sces$annotation)

(bb <- ggplot(junctions2,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), y = perc_expressed, fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab(""))

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"barplot4_TSO.svg"), plot=bb, width=4.1, height=1.9)

# Same for WTA 

junctions2<-b_wta %>% dplyr::filter(grepl("roi_1[3456]", region))
regions2<-c("[roi_13] chr5:75180272-75180999","[roi_14] chr5:75181109-75181522","[roi_15] chr5:75181675-75182976","[roi_16] chr5:75183142-75185514") # filtering canonical regions
junctions2<-junctions2[which(junctions2$region %in% regions2),]

b_wta$region_name<-gsub(" chr5","",b_wta$region_name)

names(myscale)<-unique(sces$annotation)

(bb <- ggplot(junctions2,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), y = perc_expressed, fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab(""))

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"barplot4_WTA.svg"), plot=bb, width=4.1, height=1.9)

```

### Same but for Pdgfra positive cells only 

### eGFP 

```{r,fig.width=2.3,fig.height=2.3}

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","low","pdgfra_unimodal")

datadir <- file.path(here("01_pdgfra_experiment"), 
                     "data")

# read in SummarizedExperiment with the junction counts

rm(b_tso,b_wta)

b_tso <- readRDS(file = file.path(datadir,
                              "junctions", "pdgfra_only_tso_df-annotated-junctions.rds"))

b_wta <- readRDS(file = file.path(datadir,
                              "junctions", "pdgfra_only_wta_df-annotated-junctions.rds"))

# Same plots for TSO and WTA separately, start with TSO

(aa <- ggplot(b_tso %>% dplyr::filter(grepl("egfp", region)),
                    aes(x = celltype, y = perc_expressed,
                        fill= factor(sample,levels=c("unmod","rockroi")),
                        label = perc_expressed,
                        colour = factor(sample,levels=c("unmod","rockroi"))))+
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), 
            colour = "black",
            position = position_dodge(width = 0.9),
            size=2,
            vjust=17) +
  facet_wrap(~region, nrow = 1) +
  theme_bw() +
  rock_roi_theme_multiple_panels() +
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  scale_color_manual(labels=sample,
                    values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
  theme(legend.position = "none"))+
  xlab("")

# same for WTA --> 0 

# (aa <- ggplot(b_wta[which(b_wta$region_name=="[roi_egfp] egfp"),],
#                     aes(x = celltype, y = perc_expressed,
#                         fill= factor(sample,levels=c("unmod","rockroi")),
#                         label = perc_expressed,
#                         colour = factor(sample,levels=c("unmod","rockroi"))))+
#   geom_bar(position = position_dodge2(), stat = "identity") +
#   geom_text(aes(label = round(perc_expressed, 1), 
#                 y = perc_expressed+1.5), 
#             colour = "black",
#             position = position_dodge(width = 0.9),
#             size=2,
#             vjust=17) +
#   facet_wrap(~region, nrow = 1) +
#   theme_bw() +
#   rock_roi_theme_multiple_panels() +
#   theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
#   ylab("Percent cells expressed")+
#   scale_fill_manual(labels=sample,
#                     values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
#   scale_color_manual(labels=sample,
#                     values=c(myscale[["unmod"]],myscale[["rockroi"]]))+
#   theme(legend.position = "none"))+
#   xlab("")

```

### Regions 1

```{r,fig.height=2.3,fig.width=4}

# For TSO

b_tso$region_name<-gsub(" chr5","",b_tso$region_name)

junctions1<-b_tso %>% dplyr::filter(grepl("roi_[456]", region))
regions<-c("[roi_4] chr5:75166771-75167837","[roi_5] chr5:75167967-75170495","[roi_6] chr5:75170666-75170762") # filtering canonical regions
junctions1<-junctions1[which(junctions1$region %in% regions),]

names(myscale)<-unique(sces$annotation)

bb <- ggplot(junctions1,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), 
                        y = perc_expressed, 
                        fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab("")

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"pdgfra_only_barplot3_tso.svg"), plot=bb, width=4, height=1.9)


# For WTA

b_wta$region_name<-gsub(" chr5","",b_wta$region_name)

junctions1<-b_wta %>% dplyr::filter(grepl("roi_[456]", region))
regions<-c("[roi_4] chr5:75166771-75167837","[roi_5] chr5:75167967-75170495","[roi_6] chr5:75170666-75170762") # filtering canonical regions
junctions1<-junctions1[which(junctions1$region %in% regions),]

names(myscale)<-unique(sces$annotation)

bb <- ggplot(junctions1,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), 
                        y = perc_expressed, 
                        fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions1$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab("")

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"pdgfra_only_barplot3_wta.svg"), plot=bb, width=4, height=1.9)

```

### Other regions

```{r,fig.height=2.3,fig.width=4}

# For TSO

junctions2<-b_tso %>% dplyr::filter(grepl("roi_1[3456]", region))
regions2<-c("[roi_13] chr5:75180272-75180999","[roi_14] chr5:75181109-75181522","[roi_15] chr5:75181675-75182976","[roi_16] chr5:75183142-75185514") # filtering canonical regions
junctions2<-junctions2[which(junctions2$region %in% regions2),]

b_tso$region_name<-gsub(" chr5","",b_tso$region_name)

names(myscale)<-unique(sces$annotation)

(bb <- ggplot(junctions2,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), y = perc_expressed, fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab(""))

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"pdgfra_only_barplot4_TSO.svg"), plot=bb, width=4.1, height=1.9)

# Same for WTA 

junctions2<-b_wta %>% dplyr::filter(grepl("roi_1[3456]", region))
regions2<-c("[roi_13] chr5:75180272-75180999","[roi_14] chr5:75181109-75181522","[roi_15] chr5:75181675-75182976","[roi_16] chr5:75183142-75185514") # filtering canonical regions
junctions2<-junctions2[which(junctions2$region %in% regions2),]

b_wta$region_name<-gsub(" chr5","",b_wta$region_name)

names(myscale)<-unique(sces$annotation)

(bb <- ggplot(junctions2,
                    aes(x = factor(sample,levels=c("unmod","rockroi")), y = perc_expressed, fill = celltype,
                        colour = celltype, group = celltype,
                        label = perc_expressed)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label = round(perc_expressed, 1), 
                y = perc_expressed+1.5), colour = "black",
            position = position_dodge2(width = 1),
            size=2,
            vjust=-0.2) +
  facet_wrap(~region_name, nrow = 1) +
  theme_classic()+
  theme(axis.text = element_text(size=5),
                plot.title=element_text(size=7,face="bold",hjust=0.5),
        axis.title=element_text(size=7,face="bold"),
        strip.text=element_text(size=5,face="bold"),
        legend.title=element_text(size=7,face="bold"),
        legend.text=element_text(size=5),
        legend.key.size = unit(0.5, 'cm'),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5),
        legend.title.align=0.5)+
  theme(axis.text.x = element_text(angle = 30, vjust = 1, hjust=1)) +
  ylab("Percent cells expressed")+
  scale_fill_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  scale_color_manual(labels=unique(junctions2$celltype),
                    values=c(myscale[["Crypt_bottom_1"]],myscale[["Crypt_top"]],myscale[["Enteroendocrine"]]))+
  xlab(""))

bb<-ggrastr::rasterize(bb, layers='Point', dpi=1000)

bb

ggsave(file=file.path(datadir,"pdgfra_only_barplot4_WTA.svg"), plot=bb, width=4.1, height=1.9)

```


### Sashimis by cell type

```{r,fig.height=3,fig.width=8}

names(myscale) <- c("unmod","unmod_roi","rock","egfp_cds","rockroi","human","unknown","mouse","tdtom_cds","no_detection","doublets","low","pdgfra_unimodal")

options(ucscChromosomeNames=FALSE)

anno <- rtracklayer::import(file.path(datadir, "subset.gtf")) %>% 
  as.data.frame %>% dplyr::filter(type=="exon")
colnames(anno)[1] <- "chromosome"

gtf <- rtracklayer::import("/Users/gimoro/figure_making_paper/rock_roi_paper/01_pdgfra_experiment/data/pdgfra.gtf")

gtf_path<-dir(here("01_pdgfra_experiment"), ".*gtf", 
            recursive = TRUE, full.names = TRUE)

track<-makeTxDbFromGFF(file="/Users/gimoro/figure_making_paper/rock_roi_paper/01_pdgfra_experiment/data/pdgfra.gtf", format="gtf")

region_start<-min(start(ranges(gtf[which(gtf$gene_name=="Pdgfra"),])))-100 # add extra 100 because otherwise don't see labels in plot
region_end<-max(end(ranges(gtf[which(gtf$gene_name=="Pdgfra"),])))+100

print(region_start)
print(region_end)

# annotation track Pdgfra

gm_pdgfra <- GeneRegionTrack(track,
                      name="Pdgfra",
                      geneSymbol=TRUE,
                      showId=TRUE,
                      chromosome="chr5",
                      from=region_start,
                      to=region_end,
                      transcriptAnnotation="symbol")

displayPars(gm_pdgfra)<-list(col="black",
                      cex.group=0.5,
                      col.line="black",
                      fontcolor.group="black")

displayPars(gm_pdgfra) <- list(fill="grey")


z<- GRanges("chr5", IRanges(region_start,region_end))
param <- ScanBamParam(which=z)

# Loading .bam files 

tso_rr_cryptb <- DataTrack(file.path(datadir, "tso_rockroi__Crypt_bottom.bam"),
                             name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,
                     col.histogram="grey",
                     fill.histogram="gray",
                     ylim=c(0,700))

wta_rr_cryptb <- DataTrack(file.path(datadir, "wta_rockroi__Crypt_bottom.bam"),
                             name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,
                     col.histogram="black",
                     fill.histogram="black",
                     ylim=c(0,700))

tso_a_rr_cryptb<-AlignmentsTrack(range=file.path(datadir, "tso_rockroi__Crypt_bottom.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_bottom")

wta_a_rr_cryptb<-AlignmentsTrack(range=file.path(datadir, "wta_rockroi__Crypt_bottom.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_bottom")

displayPars(tso_rr_cryptb) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="grey",
                               fill.histogram="gray")

displayPars(wta_rr_cryptb) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="black",
                               fill.histogram="black")

displayPars(tso_a_rr_cryptb) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

displayPars(wta_a_rr_cryptb) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

tso_rr_cryptt <- DataTrack(file.path(datadir, "tso_rockroi__Crypt_top.bam"),
                             name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,col.histogram="grey",
                     fill.histogram="gray",
                     ylim=c(0,200))

wta_rr_cryptt <- DataTrack(file.path(datadir, "wta_rockroi__Crypt_top.bam"),
                             name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,col.histogram="black",
                     fill.histogram="black",
                     ylim=c(0,200))

tso_a_rr_cryptt<-AlignmentsTrack(range=file.path(datadir, "tso_rockroi__Crypt_top.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_top")

wta_a_rr_cryptt<-AlignmentsTrack(range=file.path(datadir, "wta_rockroi__Crypt_top.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_top")

displayPars(tso_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="grey",
                               fill.histogram="gray")

displayPars(wta_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="black",
                               fill.histogram="black")

displayPars(wta_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

displayPars(tso_a_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

displayPars(wta_a_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

tso_rr_epi <- DataTrack(file.path(datadir, "tso_rockroi__Epithelial.bam"),
                             name = "Epithelial",
                             isPaired = FALSE,
                    chromosome="chr5",
                     from=region_start,
                     to=region_end,
                    col.histogram="grey",
                    fill.histogram="gray")

wta_rr_epi <- DataTrack(file.path(datadir, "wta_rockroi__Epithelial.bam"),
                             name = "Epithelial",
                             isPaired = FALSE,
                    chromosome="chr5",
                     from=region_start,
                     to=region_end,
                    col.histogram="black",
                    fill.histogram="black")

tso_a_rr_epi<-AlignmentsTrack(range=file.path(datadir, "tso_rockroi__Epithelial.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Epithelial")

wta_a_rr_epi<-AlignmentsTrack(range=file.path(datadir, "wta_rockroi__Epithelial.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Epithelial")

displayPars(tso_rr_epi)<-list("background.title"=myscale[["rockroi"]],
                               col.histogram="grey",
                              fill.histogram="gray")

displayPars(wta_rr_epi)<-list("background.title"=myscale[["rockroi"]],
                               col.histogram="black",
                              fill.histogram="black")

displayPars(tso_a_rr_epi) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

displayPars(wta_a_rr_epi) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

tso_um_cryptb <- DataTrack(file.path(datadir, "tso_unmod__Crypt_bottom.bam"),
                             ylim = c(0,20), name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,
                     col.histogram="grey",
                     fill.histogram="gray",
                     ylim=c(0,700))

wta_um_cryptb <- DataTrack(file.path(datadir, "wta_unmod__Crypt_bottom.bam"),
                             ylim = c(0,20), name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,
                     col.histogram="black",
                     fill.histogram="black",
                     ylim=c(0,700))

tso_a_um_cryptb<-AlignmentsTrack(range=file.path(datadir, "tso_unmod__Crypt_bottom.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_bottom")

wta_a_um_cryptb<-AlignmentsTrack(range=file.path(datadir, "wta_unmod__Crypt_bottom.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_bottom")

displayPars(tso_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               col.histogram="grey",
                               fill.histogram="gray")

displayPars(wta_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               col.histogram="black",
                               fill.histogram="black")

displayPars(tso_a_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               "col"="black")

displayPars(wta_a_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               "col"="black")

tso_um_cryptt <- DataTrack(file.path(datadir, "tso_unmod__Crypt_top.bam"),
                             ylim = c(0,20), name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,
                     col.histogram="grey",
                     fill.histogram="gray",
                     ylim=c(0,200))

wta_um_cryptt <- DataTrack(file.path(datadir, "wta_unmod__Crypt_top.bam"),
                             ylim = c(0,20), name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="chr5",
                     from=region_start,
                     to=region_end,
                     col.histogram="black",
                     fill.histogram="black",
                     ylim=c(0,200))

tso_a_um_cryptt<-AlignmentsTrack(range=file.path(datadir, "tso_unmod__Crypt_top.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_top")

wta_a_um_cryptt<-AlignmentsTrack(range=file.path(datadir, "wta_unmod__Crypt_top.bam"), 
                             chromosome="chr5",
                             from=region_start,
                             to=region_end,
                             name="Crypt_top")

displayPars(tso_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               col.histogram="grey",
                               fill.histogram="grey")

displayPars(wta_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               "col"="black",
                               col.histogram="black",
                               fill.histogram="black")

displayPars(tso_a_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               "col"="black")

displayPars(wta_a_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               "col"="black")

displayPars(wta_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               "col"="black",
                               col.histogram="black",
                               fill.histogram="black")


# overlaying tracks

rr_cryptb<-OverlayTrack(trackList=list(wta_rr_cryptb,tso_rr_cryptb))
rr_cryptt<-OverlayTrack(trackList=list(wta_rr_cryptt,tso_rr_cryptt))

displayPars(rr_cryptb)<-list(background.title=myscale[["rockroi"]])
displayPars(rr_cryptt)<-list(background.title=myscale[["rockroi"]])
    
type<-c("histogram","sashimi")

# plotTracks(c(tso_rr_cryptb, 
#              tso_a_rr_cryptb,
#              tso_um_cryptb,
#              tso_a_um_cryptb,
#              tso_rr_cryptt, 
#              tso_a_rr_cryptt,
#              tso_um_cryptt,
#              tso_a_um_cryptt,
#              gm_pdgfra),
#            from = region_start, 
#            to = region_end,
#            type=rep(type,6),
#            chromosome = "chr5",
#            sashimiNumbers=TRUE, 
#            col.axis="black",
#            fontcolor.title="black",
#            cex.axis = 0.5,
#            cex.title=0.5,
#            lwd.title=0.5,
#            sashimiHeight=2,
#            cex=0.5,
#            minSashimiHeight=0,
#            sashimiScore=10,
#            sizes=c(1,2,1,0.1,1,2,1,0.1,2),
#            lwd.sashimiMax=2)

plotTracks(c(rr_cryptb, 
             wta_um_cryptb,
             rr_cryptt, 
             wta_um_cryptt,
             gm_pdgfra),
           from = region_start, 
           to = region_end,
           type=rep("histogram",6),
           chromosome = "chr5",
           sashimiNumbers=TRUE, 
           col.axis="black",
           fontcolor.title="black",
           cex.axis = 0.5,
           cex.title=0.5,
           lwd.title=0.5,
           sashimiHeight=2,
           cex=0.5,
           minSashimiHeight=0,
           sashimiScore=10,
           sizes=c(1,1,1,1,1),
           lwd.sashimiMax=2)



```

### Representation zoomed in region Pdgfra

```{r,fig.heigh=40,fig.width=4}

type<-c("histogram","pileup","sashimi")

print(region_start+30000)
print(region_end-12500)

plotTracks(c(rr_cryptb, 
             tso_a_rr_cryptb,
             gm_pdgfra),
           from = region_start+30000, 
           to = region_end-12500,
           type=type,
           chromosome = "chr5",
           sashimiNumbers=TRUE, 
           col.axis="black",
           fontcolor.title="black",
           cex.axis=0.5,
           cex.title=0.5,
           lwd.title=0.5,
           cex=0.5,
           col.gaps="orange",
           lwd.sashimiMax=2,
           lwd.gaps=0.6,
           stacking="squish",
           sashimiScore=1,
           max.height=0.7,
           min.height=0.7,
           sashimiHeight=2,
           sizes=c(3,30,3),
           ylim=c(0,300))

plotTracks(c(wta_um_cryptb, 
             wta_a_um_cryptb,
             gm_pdgfra),
           from = region_start+30000, 
           to = region_end-12500,
           type=type,
           chromosome = "chr5",
           sashimiNumbers=TRUE, 
           col.axis="black",
           fontcolor.title="black",
           cex.axis=0.5,
           cex.title=0.7,
           lwd.title=0.7,
           cex=0.5,
           col.gaps="orange",
           lwd.sashimiMax=2,
           lwd.gaps=0.6,
           stacking="squish",
           sashimiScore=1,
           max.height=0.5,
           min.height=0.5,
           sashimiHeight=2,
           sizes=c(3,30,3),
           ylim=c(0,150))

plotTracks(c(rr_cryptt, 
             tso_a_rr_cryptt,
             gm_pdgfra),
           from = region_start+30000, 
           to = region_end-12500,
           type=type,
           chromosome = "chr5",
           sashimiNumbers=TRUE, 
           col.axis="black",
           fontcolor.title="black",
           cex.axis=0.5,
           cex.title=0.5,
           lwd.title=0.5,
           cex=0.5,
           col.gaps="orange",
           lwd.sashimiMax=2,
           lwd.gaps=0.6,
           stacking="squish",
           sashimiScore=1,
           max.height=0.7,
           min.height=0.7,
           sashimiHeight=2,
           sizes=c(3,30,3),
           ylim=c(0,40))

plotTracks(c(wta_um_cryptt, 
             wta_a_um_cryptt,
             gm_pdgfra),
           from = region_start+30000, 
           to = region_end-12500,
           type=type,
           chromosome = "chr5",
           sashimiNumbers=TRUE, 
           col.axis="black",
           fontcolor.title="black",
           cex.axis=0.5,
           cex.title=0.7,
           lwd.title=0.7,
           sashimiHeight=2,
           cex=0.5,
           col.gaps="orange",
           lwd.sashimiMax=2,
           lwd.gaps=0.6,
           stacking="squish",
           sashimiScore=1,
           max.height=0.5,
           min.height=0.5,
           sizes=c(3,30,3),
           ylim=c(0,70))

```

### eGFP 

```{r,fig.width=2.7,fig.height=2.7}

options(ucscChromosomeNames=FALSE)

anno <- rtracklayer::import(file.path(datadir, "subset.gtf")) %>% 
  as.data.frame %>% dplyr::filter(type=="exon")
colnames(anno)[1] <- "chromosome"

gtf <- rtracklayer::import("/Users/gimoro/figure_making_paper/rock_roi_paper/01_pdgfra_experiment/data/pdgfra.gtf")

gtf_path<-dir(here("01_pdgfra_experiment"), ".*gtf", 
            recursive = TRUE, full.names = TRUE)

track<-makeTxDbFromGFF(file="/Users/gimoro/figure_making_paper/rock_roi_paper/01_pdgfra_experiment/data/pdgfra.gtf", format="gtf")

region_start<-min(start(ranges(gtf[which(seqnames(gtf)=="egfp"),])))-100 # add extra 10000 because otherwise don't see labels in plot
region_end<-max(end(ranges(gtf[which(seqnames(gtf)=="egfp"),])))

print(region_start)
print(region_end)

gm_egfp <- GeneRegionTrack(track,
                      name="egfp",
                      geneSymbol=TRUE,
                      showId=TRUE,
                      chromosome="egfp",
                      from=region_start,
                      to=region_end,
                      transcriptAnnotation="symbol")

displayPars(gm_egfp)<-list(col="black",
                      cex.group=0.5,
                      col.line="black",
                      fontcolor.group="black")

displayPars(gm_egfp) <- list(fill= "grey")

z<- GRanges("egfp", IRanges(region_start,region_end))
param <- ScanBamParam(which=z)

tso_rr_cryptb <- DataTrack(file.path(datadir, "tso_rockroi__Crypt_bottom.bam"),
                             name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,
                     col.histogram="grey",
                     fill.histogram="gray")

wta_rr_cryptb <- DataTrack(file.path(datadir, "wta_rockroi__Crypt_bottom.bam"),
                             name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,
                     col.histogram="black",
                     fill.histogram="black")

displayPars(tso_rr_cryptb) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="grey",
                               fill.histogram="gray")

displayPars(wta_rr_cryptb) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="black",
                               fill.histogram="black")

tso_rr_cryptt <- DataTrack(file.path(datadir, "tso_rockroi__Crypt_top.bam"),
                             name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,col.histogram="grey",
                     fill.histogram="gray")

wta_rr_cryptt <- DataTrack(file.path(datadir, "wta_rockroi__Crypt_top.bam"),
                             name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,col.histogram="black",
                     fill.histogram="black")

displayPars(tso_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="grey",
                               fill.histogram="gray")

displayPars(wta_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               col.histogram="black",
                               fill.histogram="black")

displayPars(wta_rr_cryptt) <- list("background.title"=myscale[["rockroi"]],
                               "col"="black")

tso_um_cryptb <- DataTrack(file.path(datadir, "tso_unmod__Crypt_bottom.bam"),
                             ylim = c(0,20), name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,
                     col.histogram="grey",
                     fill.histogram="gray")

wta_um_cryptb <- DataTrack(file.path(datadir, "wta_unmod__Crypt_bottom.bam"),
                             ylim = c(0,20), name = "Crypt_bottom",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,
                     col.histogram="black",
                     fill.histogram="black")

tso_a_um_cryptb<-AlignmentsTrack(range=file.path(datadir, "tso_unmod__Crypt_bottom.bam"), 
                             chromosome="egfp",
                             from=region_start,
                             to=region_end,
                             name="Crypt_bottom")

wta_a_um_cryptb<-AlignmentsTrack(range=file.path(datadir, "wta_unmod__Crypt_bottom.bam"), 
                             chromosome="egfp",
                             from=region_start,
                             to=region_end,
                             name="Crypt_bottom")

displayPars(tso_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               col.histogram="grey",
                               fill.histogram="gray")

displayPars(wta_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               col.histogram="black",
                               fill.histogram="black")

displayPars(tso_a_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               "col"="black")

displayPars(wta_a_um_cryptb) <- list("background.title"=myscale[["unmod"]],
                               "col"="black")

tso_um_cryptt <- DataTrack(file.path(datadir, "tso_unmod__Crypt_top.bam"),
                              name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,
                     col.histogram="grey",
                     fill.histogram="gray")

wta_um_cryptt <- DataTrack(file.path(datadir, "wta_unmod__Crypt_top.bam"),
                              name = "Crypt_top",
                             isPaired = FALSE,
                       chromosome="egfp",
                     from=region_start,
                     to=region_end,
                     col.histogram="black",
                     fill.histogram="black")

displayPars(tso_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               col.histogram="grey",
                               fill.histogram="grey")

displayPars(wta_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               "col"="black",
                               col.histogram="black",
                               fill.histogram="black")

displayPars(wta_um_cryptt) <- list("background.title"=myscale[["unmod"]],
                               "col"="black",
                               col.histogram="black",
                               fill.histogram="black")


# overlaying tracks

rr_cryptb<-OverlayTrack(trackList=list(wta_rr_cryptb,tso_rr_cryptb))
rr_cryptt<-OverlayTrack(trackList=list(wta_rr_cryptt,tso_rr_cryptt))

displayPars(rr_cryptb)<-list(background.title=myscale[["rockroi"]])
displayPars(rr_cryptt)<-list(background.title=myscale[["rockroi"]])
    
type<-c("histogram","sashimi")

plotTracks(c(rr_cryptb, 
             wta_um_cryptb,
             rr_cryptt, 
             wta_um_cryptt,
             gm_egfp),
           from = region_start, 
           to = region_end,
           type="histogram",
           chromosome = "egfp",
           sashimiNumbers=TRUE, 
           col.axis="black",
           fontcolor.title="black",
           cex.axis = 0.5,
           cex.title=0.5,
           lwd.title=0.5,
           cex=0.5,
           sashimiHeight=20, 
           minSashimiHeight=0,
           sizes=c(1,1,1,1,1),
           lwd.sashimiMax=2,
           ylim=c(0,800))

```

### Session info

```{r}

sessionInfo()

```

